\chapter{Conclusion}

The conclusion summarizes the overall work carried out in the project and highlights the major achievements, findings, and outcomes obtained during the development and evaluation phases. This chapter discusses the effectiveness of the proposed system in addressing the identified problem statement and achieving the specified objectives. The limitations of the current implementation and possible future enhancements are also presented to provide direction for further research and development.

%----------------------------------------------------------

\section{Limitations}

Although the proposed system demonstrates effective performance and functionality, certain limitations may still exist due to technical, operational, or resource-related constraints. This section discusses the boundaries and restrictions of the current implementation.

The limitations may include:
\begin{itemize}
    \item Limited dataset availability or size
    \item Hardware and computational constraints
    \item Dependency on internet connectivity or external services
    \item Scalability limitations for very large-scale deployments
    \item Reduced accuracy under specific conditions
    \item Security and privacy considerations
\end{itemize}

The discussion should clearly specify the conditions under which the proposed system performs effectively and the scenarios where improvements may be required.

%----------------------------------------------------------

\section{Future Enhancement}

This section discusses the possible improvements and additional functionalities that can be incorporated into the system in future versions. Future enhancements help improve scalability, performance, usability, and adaptability of the proposed solution.

Potential future enhancements may include:
\begin{itemize}
    \item Integration with cloud computing platforms
    \item Real-time data processing and analytics
    \item Mobile application support
    \item Enhanced security and authentication mechanisms
    \item Artificial Intelligence (AI) and Machine Learning (ML) based automation
    \item Improved scalability and distributed processing capabilities
    \item User interface and visualization improvements
\end{itemize}

Future work may also focus on optimizing algorithms, improving system efficiency, and expanding the application scope of the project.

%----------------------------------------------------------

\vspace{1cm}

The successful implementation and evaluation of the proposed system demonstrate its practical applicability and effectiveness in addressing the identified problem domain. The methodologies, architectures, and technologies adopted during the project provide a strong foundation for future research and further development in this area.