%%ecproject package is created by P Narashimaraja, Assistant Professor, ECE, RVCE.
%%Border
\begin{tikzpicture}[remember picture, overlay]
	\draw[line width = 4pt] ($(current page.north west) + (0.75in,-0.75in)$) rectangle ($(current page.south east) + (-0.75in,0.75in)$);
\end{tikzpicture}
\thispagestyle{empty}
\ifIDP \vspace{-1cm}\fi
\begin{center}
	\Large\textbf{\underline{ACKNOWLEDGEMENT}} \par
\end{center}

\vspace{0.2cm}

\begingroup
\setlength{\parskip}{0.5\baselineskip}

I am indebted to my guide, \textbf{Dr.~Manonmani S}, Associate Professor, Department of CSE, RVCE for the suggestions \& advice through the project work.\par

I extend my sincere gratitude to \textbf{Prof.~Dr.~Holger Schmidt} and \textbf{Prof.~Dr.~Klaus Kaiser}, Department of Computer Science, Fachhochschule Dortmund, for their invaluable guidance, and for welcoming this collaboration from RV College of Engineering.\par

My sincere thanks to the project coordinators \textbf{Dr.~Chethana R Murthy}, Associate Professor, Department of CSE and \textbf{Prof.~Deepika Dash}, Assistant Professor, Department of CSE for their timely instructions and support in coordinating the project.\par

My gratitude to \textbf{Prof.~Narashimaraja P}, Department of ECE, RVCE for the organized latex template which made report writing easy.\par

My sincere thanks to \textbf{Dr.~Shanta Rangaswamy}, Professor and Head, Department of Computer Science Engineering, RVCE and \textbf{Dr.~Ramakanth Kumar P}, Professor and Dean (CSE Cluster), RVCE for the support and encouragement.\par

I express sincere gratitude to our beloved \textbf{Dr.~M.V. Renukadevi}, Professor and Dean Academics, RVCE, \textbf{Dr.~Geetha K S}, Professor and Vice Principal, RVCE and \textbf{Dr.~K. N. Subramanya}, Principal, RVCE for the appreciation towards this project work.\par

I thank all the teaching staff and technical staff of Computer Science Engineering department, RVCE for their help. Lastly, I take this opportunity to thank my family members and friends who provided all the backup support throughout the project work.\par

\endgroup

\vspace{1cm}
\begin{flushright}
\textbf{\printStuNameA}\\ \vspace{-2pt}
\ifIDP \printGuideDeptAInLF\else\printDepartmentLF\fi, \\ \vspace{-2pt}
RV College of Engineering, \\ \vspace{-2pt}
Bengaluru-59.
\end{flushright}
