\section{Conclusion}
\label{sec:conclusion}

Most Python vulnerability detectors today are evaluated on labels
inherited unmodified from CVE fix commits, on metrics that do not
account for the rarity of vulnerable code, and on a single point on
the input distribution. Each of these is a substantive source of
measurement noise on its own, and together they produce comparisons
that are difficult to interpret and harder to replicate. The CASTLE
benchmark~\cite{castle} addressed two of these issues for C and
C++ with a hand-curated 250-program corpus, but its design does not
transfer to Python, where the dominant CWE classes are sink-anchored
injection and deserialization weaknesses rather than memory-safety
bugs, and where the labels themselves are noisier because Python
advisories rarely localize the defect below the file.

This paper contributes a Python-specific evaluation substrate that
addresses the three issues directly. The benchmark covers seven
sink-shaped MITRE Top-25 CWE classes with 871 positives and 429 safe
hard negatives, partitioned into 607 training, 127 validation, and
132 test samples under a repo-group stratified split with zero
cross-split repository leakage. Labels are validated through a
three-layer methodology that combines a sink-presence filter, three
rounds of independent adversarial audit, and a diff-based filter that
requires the sink line to have been modified between
\texttt{code\_before} and \texttt{code\_after}. The combined
methodology reduces the audited false-positive rate from 52\% on raw
CVE-feed labels to 26\%. We additionally publish seven
semantics-preserving adversarial mutators across three orthogonal
families (sink-preserving: dead-code injection, string literal
splitting, identifier rename, wrapper extraction; sink-targeted:
attribute-access obfuscation, builtins-module indirection; dataflow:
taint-through-dict), each targeting a known detection shortcut, and
use them to extend a single-axis detection comparison to a two-axis
comparison of detection accuracy against rewrite robustness. To our knowledge this is the
first Python vulnerability benchmark that publishes an automated,
sink-anchored label-validity criterion and an adversarial robustness
probe alongside detection accuracy.

The current evaluation reports two static-analysis baselines
(Bandit, Semgrep) and one benchmark-trained learned detector
(a fine-tuned GraphCodeBERT classifier) on the clean test set and
the seven adversarial variants under a five-measure profile
(macro-F1, detection MCC, severity-weighted recall, hierarchical
CWE macro-accuracy, weighted Cohen's $\kappa$). The headline finding
is that the sink-token mutator family separates the detector tiers
cleanly: the two SAST tools drop 0.11--0.19 macro-F1 and
0.20--0.35 severity-weighted recall under attribute-access
obfuscation and bare-builtin indirection, while the trained detector
is invariant to every mutator in the suite. Three follow-up tracks
are planned. First, an LLM detector row (Anthropic Claude) is
built into the harness and reserved for an authorized run. Second,
the multi-model learned-detector evaluation (a second Transformer
encoder and a non-Transformer baseline) is the natural extension to
generalize the single-checkpoint robustness claim. Third, the
deferred structural CWE track (CWE-352, CWE-862, CWE-863, CWE-306,
CWE-284, CWE-639, CWE-200) will be evaluated separately because the
sink-anchored detection methodology cannot reliably detect the
absence-of-a-check vulnerability shape; a structural-CWE benchmark
requires a different methodology built around application-level
context that we leave to follow-up work.

\paragraph{Data and code availability}
The benchmark dataset, the audit reports for all three rounds, the
rejection manifests with per-sample reason codes for every dropped
sample, the regex pattern set, the diff-filter and mutator
implementations, the audit-pack generator, and the evaluation harness
are released under a permissive open-source license. The dataset is
hash-pinned to a tagged release. Any future revision will be a
separate tag with a documented changelog. % TODO: insert repository URL
% and dataset DOI (Zenodo / figshare) before submission.

\paragraph{Acknowledgements}
% TODO: acknowledge supervisors, funding source (if applicable), and
% compute provider (if applicable). Funding statements may need to
% appear in a separate \section*{Funding} block per Elsevier policy.
