\section{Dataset Construction}
\label{sec:dataset}

This section describes the substrate of the benchmark: how samples are
selected, which sources they come from, and how labels are validated.
The three-layer label validation methodology
(Sections~\ref{sec:dataset:sink},~\ref{sec:dataset:audit},
and~\ref{sec:dataset:diff}) is the principal methodological
contribution; the final composition (Section~\ref{sec:dataset:final})
is the artifact that downstream evaluation (Section~\ref{sec:evaluation})
operates on.

\subsection{Scope and CWE selection}
\label{sec:dataset:scope}

The MITRE Top-25 Most Dangerous Software Weaknesses~\cite{TODO_mitre_top25}
lists 25 CWE classes by aggregate severity-weighted exploit prevalence
across all NVD-tracked CVEs. Sixteen of the 25 are
\emph{Python-relevant}: nine entries describe C/C++ memory bugs
(CWE-787, CWE-125, CWE-416, CWE-476, CWE-119, CWE-190, CWE-120, CWE-121,
CWE-122) that surface only through CPython's C extensions, where the
bug lives in the underlying C code, and are out of scope.

We further partition the sixteen Python-relevant CWEs into two
methodological categories.

\paragraph{Sink-shaped CWEs} have a well-defined unsafe function or
construct (a ``sink'') whose unsafe use is the vulnerability. The fix
modifies the sink call, typically replacing it with one of a small set
of safe alternatives:
\begin{itemize}
\item \textbf{CWE-89 (SQL Injection)}: \texttt{cursor.execute(f"...")}
  fixed to \texttt{cursor.execute("...?", (val,))}.
\item \textbf{CWE-79 (XSS)}: \texttt{Markup(user\_input)} fixed to
  \texttt{escape(user\_input)} or removed.
\item \textbf{CWE-22 (Path Traversal)}: \texttt{open(taint)} fixed to
  \texttt{open(safe\_join(BASE, secure\_filename(taint)))}.
\item \textbf{CWE-78 (OS Command Injection)}: \texttt{os.system(f"...
  \{taint\}")} fixed to \texttt{subprocess.run([...,~taint],
  shell=False)}.
\item \textbf{CWE-94 (Code Injection)}: \texttt{eval(taint)} removed
  or replaced with \texttt{ast.literal\_eval}.
\item \textbf{CWE-918 (SSRF)}: \texttt{requests.get(url)} restricted
  with hostname allowlists or \texttt{is\_private\_address} checks.
\item \textbf{CWE-502 (Insecure Deserialization)}:
  \texttt{pickle.loads(taint)} replaced with
  \texttt{yaml.safe\_load} or signed payloads.
\end{itemize}

Two additional sink-shaped CWEs were initially included and later
removed during cleanup (see Section~\ref{sec:dataset:audit}):
CWE-434 (Unrestricted File Upload), where audit revealed sink patterns
predominantly matched framework source rather than application
handlers; and CWE-798 (Hardcoded Credentials), where all samples came
from a single static miner (single-source bias) and the
two-bucket fix alphabet did not align with the multi-variant
``alphabet'' framing applied to the other classes. CWE-77 (Improper
Neutralization in a Command) was merged into CWE-78 (its child in the
MITRE hierarchy) because in Python the two CWEs fire on the same sink
set; preserving them as separate labels created artificial scarcity.

\paragraph{Structural CWEs} encode the \emph{absence} of a check
rather than the misuse of a sink. The vulnerability and its fix
differ only in whether a defensive construct is present:
\begin{itemize}
\item CWE-352 (CSRF): the absent construct is a CSRF token validator
  or middleware.
\item CWE-862 / CWE-863 (Missing / Incorrect Authorization): the
  absent construct is an authorization check
  (\texttt{@permission\_required} decorator, ownership filter, etc.).
\item CWE-306 (Missing Authentication): the absent construct is
  \texttt{@login\_required} on a privileged route.
\item CWE-284 (Improper Access Control): umbrella for the above.
\item CWE-639 (IDOR): the absent construct is a per-request
  ownership filter on a queried object.
\item CWE-200 (Information Exposure): partially sink-shaped (output
  channels) but the vulnerable cases require knowing which fields
  are sensitive in context.
\end{itemize}

The sink-anchored detection methodology that this benchmark evaluates
--- regex sink patterns, file-level static analysis, and code-as-text
LLM prompts --- cannot reliably detect structural CWEs without
application-level context. We exclude structural CWEs from the
benchmark and document the analysis as scope justification rather
than a limitation. CWE-20 (Improper Input Validation) and CWE-770
(Allocation Without Limits) are also excluded as too broad to admit
a closed fix-pattern alphabet.

\paragraph{Final scope.}
The benchmark covers \textbf{seven sink-shaped Python CWE classes}
(CWE-89, CWE-79, CWE-22, CWE-78, CWE-94, CWE-918, CWE-502) plus a
\texttt{safe} hard-negative class for negative evaluation, for a
total of eight labels. This figure sits comfortably under the
typical 10-label cap for stable multi-class
evaluation~\cite{TODO_classification_eval_caps}.

\subsection{Data sources}
\label{sec:dataset:sources}

Samples are aggregated from five primary sources, supplemented by
two derived sources (hard negatives and hand-curated canonical
samples). Table~\ref{tab:sources} summarizes each source's
retrieval method, raw yield, and the fraction kept after the
three label-validation layers.

\begin{table}[h]
\centering
\caption{Data sources and per-source yield after label validation.
Counts reflect the final benchmark composition.}
\label{tab:sources}
\begin{tabular}{lrl}
\toprule
\textbf{Source} & \textbf{Final samples} & \textbf{Retrieval method} \\
\midrule
GHSA Advisory DB    & 384 & Local clone (1.5\,GB), per-advisory commit fetch via GitHub API \\
\texttt{vudenc}     & 209 & Pre-labeled corpus (no CVE pair) \\
CVEfixes            &  58 & Local SQLite DB (23\,GB) query on \texttt{file\_change} \\
NVD-targeted        &  33 & NVD API for selected CWE + Python keyword filter \\
OSV / GHSA GraphQL  &  54 & OSV API + GHSA GraphQL for commit references \\
Canonical (curated) &  20 & Hand-written textbook OWASP/SANS examples \\
\addlinespace
Hard negatives      & 429 & Per-source: paired safe variants from each scrape \\
\bottomrule
\end{tabular}
\end{table}

GHSA Advisory DB is the dominant source after Phase~2B re-scoping
because it indexes per-CWE advisories with explicit fix commits and
\texttt{database\_specific.cwe\_ids}, allowing precise per-CWE
scraping with low retrieval overhead. The \texttt{vudenc}
corpus~\cite{TODO_vudenc} is a pre-labeled human-annotated dataset
without paired before/after versions; it contributes only to CWE-89
(its strongest class by curator agreement) and is treated as a
separate label-confidence tier (Section~\ref{sec:dataset:final}).

Hand-curated canonical samples (twenty in total, distributed
across CWE-502, CWE-94, CWE-78, and CWE-434 prior to CWE-434
exclusion) are textbook OWASP and SANS exemplars rewritten as
self-contained Python files. They serve as benchmark calibration
points: a detector that misses these samples is broken regardless
of its CVE-corpus performance.

\subsection{Sample schema}
\label{sec:dataset:schema}

Each sample is stored as a paired \texttt{\{id\}.py} (the
\texttt{code\_before} source) and \texttt{\{id\}.meta.json} (a
schema-versioned metadata record). The metadata schema contains 46
fields organized into eight groups: identity (4), advisory IDs (4),
vulnerability label (5), CVSS severity decomposition (13),
provenance (7), code payload (2), code-quality signals (8), and
pipeline state (3). The complete per-field literature justification
is documented separately in \texttt{FIELD\_JUSTIFICATION.md}; here
we describe the four signal fields most relevant to label
validation:

\begin{itemize}
\item \texttt{cwe} --- the labeled MITRE CWE identifier (e.g.,
  \texttt{CWE-89}). Inherited from the source advisory.
\item \texttt{has\_cwe\_sink} --- boolean, auto-computed by the
  sink-presence filter (Section~\ref{sec:dataset:sink}).
\item \texttt{sink\_pattern} --- the specific regex pattern that
  matched, preserved for audit and reproducibility.
\item \texttt{label\_source} --- one of \texttt{advisory},
  \texttt{static\_mining}, \texttt{canonical}, or
  \texttt{commit\_classifier}. Tracks the provenance of the CWE
  assignment for label-confidence tiering.
\end{itemize}

\subsection{Layer 1: Pre-ingest sink-presence filter}
\label{sec:dataset:sink}

The first validation layer rejects samples that lack the
category-defining sink token for their labeled CWE. For each CWE
$c \in \text{TARGET\_CWES}$, we define a small set of regex sink
patterns $S_c$ corresponding to the function calls, attribute
accesses, or syntactic constructs whose unsafe use is the
vulnerability. A sample with $\texttt{cwe} = c$ passes Layer~1 if
and only if at least one pattern in $S_c$ matches its
\texttt{code\_before}. Patterns are designed for high recall: they
match \emph{both} safe and unsafe forms of the sink (Layer~3 will
discriminate them via diff comparison).

We strip line comments and triple-quoted docstrings before matching
to suppress documentation-style false positives (e.g.,
\texttt{\# pickle.loads is unsafe}). For two CWEs the sink alone is
insufficient: CWE-22 additionally requires an HTTP-request reference
within $\pm 20$ lines of the sink (test fixtures with hardcoded paths
are otherwise common false positives); CWE-798 (during its inclusion
period) required a security-context keyword within $\pm 10$ lines.

\begin{figure}[h]
\begin{verbatim}
def has_cwe_sink(code, cwe, file_path=None):
    if cwe in TEST_EXCLUSION_CWES and is_test_file(file_path):
        return False, None
    match_code = strip_comments_and_docstrings(code)
    for pat in SINK_PATTERNS[cwe]:
        for m in pat.finditer(match_code):
            if cwe in REQUIRES_REQUEST_PROXIMITY:
                if not has_request_near(code, m.start()):
                    continue
            if cwe in REQUIRES_SECURITY_CONTEXT:
                if not has_security_context_near(code, m.start()):
                    continue
            return True, pat.pattern
    return False, None
\end{verbatim}
\caption{Layer-1 sink-presence filter (simplified). Full implementation
in \texttt{src/utils/file\_utils.py:has\_cwe\_sink}.}
\label{fig:sink-filter}
\end{figure}

Application of Layer~1 to the pre-filter corpus (1{,}982 samples)
dropped 749 samples (38\%). The rejection rate varied substantially
by CWE: 68\% for CVEfixes-sourced samples (commit-level mislabels
confirmed at scale), 41\% for GHSA-DB, 28\% for NVD-targeted. The
\texttt{vudenc}-sourced CWE-94 samples were rejected wholesale via
an explicit source-CWE blocklist after audit revealed 14 of 15
samples lacked the sink set (see Section~\ref{sec:dataset:audit}).

\subsection{Layer 2: Adversarial audit}
\label{sec:dataset:audit}

The sink-presence filter rejects samples that lack a sink token but
cannot reject samples where the sink \emph{is} present but is not
the actual vulnerability (a co-changed file in a CVE fix commit).
To characterize and remove this residual noise, we conducted three
rounds of independent adversarial audit on randomly sampled
post-Layer-1 positives.

\paragraph{Methodology.}
For each audit round, the dataset is sampled stratified by CWE: ten
samples per populous class ($n > 30$), full audit for rare classes
($n \le 30$). Canonical and \texttt{safe} samples are excluded
(positive by construction, separate negative-class audit). For each
sampled sample, the auditor receives a structured prompt containing:
the sample ID, source, repository, file path, the regex pattern
that originally matched, and the $\pm 8$-line code excerpt around
the sink match. The auditor returns a binary verdict
(\textsc{pass} = correctly labeled / \textsc{fail} = mislabeled)
with a one-sentence justification. Auditors are independent and do
not see each other's verdicts.

Round~1 (77~samples, eight CWEs including CWE-434 and CWE-798)
produced an aggregate false-positive rate of 52\%. The
per-sample reasoning across the 40 \textsc{fail} verdicts
clustered into six recurring patterns, summarized in
Table~\ref{tab:audit-patterns}.

\begin{table}[h]
\centering
\caption{Audit failure patterns by frequency (round~1).}
\label{tab:audit-patterns}
\begin{tabular}{p{5cm}rp{6cm}}
\toprule
\textbf{Pattern} & \textbf{Frequency} & \textbf{Description} \\
\midrule
Sink-too-broad & 15+ & Pattern matches framework source, type annotations, imports rather than application sink call \\
Co-changed file noise & 12+ & Sink present but is parameterized / safe; fix touched the file for unrelated reasons \\
Documentation matches & 8+ & Sink pattern matches text in line comments or docstrings \\
Pattern conflation & 3 & \texttt{re.compile} matched the CWE-94 \texttt{compile(} pattern \\
Inverted operation & 1 & \texttt{jsonpickle.encode} matched the CWE-502 \texttt{jsonpickle.} pattern \\
Wrapper abstraction & 2 & Class or function wrapper of the sink, not a direct invocation \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Pattern tightenings.}
Each failure pattern motivated a specific change to either
\texttt{SINK\_PATTERNS} or the surrounding gate logic:
documentation matches were suppressed by extending the
comment-stripping pass to all CWEs; pattern conflation was
addressed by removing the bare \texttt{compile(} pattern from
CWE-94's sink set (the remaining
\texttt{eval}/\texttt{exec}/\texttt{\_\_import\_\_} patterns cover
the actual cases); inverted operation by narrowing the
\texttt{jsonpickle.} pattern to \texttt{jsonpickle.(decode|loads)}.
The CWE-22 patterns were augmented with a
\texttt{has\_request\_near} proximity check ($\pm 20$ lines)
because the unmodified sinks (\texttt{open}, \texttt{os.path.join})
match too permissively in test fixtures.

Two CWEs proved to have systematic issues that single-pattern
fixes could not address: CWE-434 (Unrestricted File Upload), where
the audit found 0 of 7 samples valid (100\% FP), and CWE-798
(Hardcoded Credentials), already flagged for the single-source-bias
issue. CWE-434 was dropped from the active taxonomy with all 15
samples (including 8 canonicals) moved to a rejection manifest; the
methodology contribution from CWE-434 remains documented in the
deferred-CWE analysis. CWE-798 was dropped at the same time.

\paragraph{Round~2 and~3.}
Round~2 (70~samples on the post-pattern-tightening dataset)
confirmed that the pattern fixes worked where targeted but
revealed an audit-floor effect: the aggregate FP rate held at 51\%,
because the new random samples surfaced different failure modes
on the unchanged-pattern CWEs. The CWEs where pattern fixes
targeted (CWE-502, CWE-22) improved; CWEs without pattern fixes
held steady. This motivated Layer~3.

Round~3 (78~samples, post-Layer-3) is reported in
Section~\ref{sec:dataset:diff}.

\subsection{Layer 3: Diff-based filter}
\label{sec:dataset:diff}

The dominant audit failure pattern across all three rounds is
``co-changed file noise'': a file touched in a CVE fix commit
whose labeled CWE is not actually localized to that file. The
sink-presence filter cannot detect this because the sink token
\emph{is} present --- it just isn't the line the fix modified.

We address this with an evidence-anchored label criterion: a
positive sample is valid only if at least one line containing a
sink-pattern match for its CWE differs between \texttt{code\_before}
and \texttt{code\_after}. Samples where every sink line is
identical between the two versions are treated as commit-level
noise. Figure~\ref{fig:diff-filter} presents the algorithm.

\begin{figure}[h]
\begin{verbatim}
def sink_was_modified(code_before, code_after, cwe):
    patterns = SINK_PATTERNS.get(cwe)
    if not patterns or not code_after.strip():
        return None  # cannot determine

    def sink_lines(code):
        out = set()
        for ln in strip_comments(code).splitlines():
            normalized = " ".join(ln.split())
            if not normalized:
                continue
            for p in patterns:
                if p.search(ln):
                    out.add(normalized)
                    break
        return out

    before = sink_lines(code_before)
    after  = sink_lines(code_after)
    return (before - after) != set()
\end{verbatim}
\caption{Diff-based filter. A sample passes if at least one sink
line in \texttt{code\_before} is absent from \texttt{code\_after}.
Comments are stripped and whitespace is normalized before
comparison so that reformat-only edits are not counted as real
changes. Implementation: \texttt{src/utils/file\_utils.py:
sink\_was\_modified}.}
\label{fig:diff-filter}
\end{figure}

\paragraph{Application.}
Of the 738 positives reaching Layer~3, 316 (43\%) had unchanged
sink lines and were rejected. The yield was strongly per-CWE:
CWE-79 lost 116 of 170 (68\%), CWE-22 lost 37 of 52 (71\%), CWE-78
lost 32 of 49 (65\%); CWE-502 lost only 18 of 90 (20\%) and
CWE-94 lost 25 of 55 (45\%). The dramatic CWE-79 cut confirms
the audit-round observation: XSS labels in CVE-fix commits
frequently mark template-related files where the fix touched
escape logic elsewhere, leaving the sink line itself unchanged.

Samples for which \texttt{code\_after} is empty (typically
\texttt{vudenc} entries, which are human-labeled without a paired
fix commit) cannot be discriminated by Layer~3. These 217 samples
are retained as a separate label-confidence tier
(Section~\ref{sec:dataset:final}).

\paragraph{Round-3 audit (post-Layer-3).}
A third audit round on 78 samples drawn from the diff-filtered
dataset measured an aggregate FP rate of 26\% --- approximately
halving the post-Layer-2 rate (Table~\ref{tab:audit-progression}).
The improvement is most pronounced on the CWEs that suffered most
from co-changed-file noise: CWE-22 dropped from 80\% to 27\% FP,
CWE-502 to 0\%, CWE-94 to 0\%. Residual noise concentrates on
CWE-89 (because \texttt{vudenc} samples are kept without diff
validation) and CWE-79 (where the diff filter does identify
\emph{some} change but the change is sometimes a defensive
no-op rather than a sink rewrite).

\begin{table}[h]
\centering
\caption{Audit FP rate progression across the three validation
layers. Each round samples disjoint subsets of the current
post-filter dataset.}
\label{tab:audit-progression}
\begin{tabular}{lrrr}
\toprule
\textbf{CWE} & \textbf{Round~1 (raw)} & \textbf{Round~2 (patterns)} & \textbf{Round~3 (+diff)} \\
\midrule
CWE-502 & 40\% & 20\% & \textbf{0\%} \\
CWE-94  & 40\% & 50\% & \textbf{0\%} \\
CWE-78  & 20\% & 30\% & 12\% \\
CWE-22  & 80\% & 70\% & 27\% \\
CWE-918 & 50\% & 70\% & 36\% \\
CWE-89  & 40\% & 30\% & 50\%$^\star$ \\
CWE-79  & 60\% & 70\% & 50\% \\
\midrule
\textbf{Overall} & \textbf{52\%} & \textbf{51\%} & \textbf{26\%} \\
\bottomrule
\end{tabular}
\\[1ex]
{\footnotesize $^\star$Round-3 CWE-89 sample was $N=6$; wide CI.
The estimate carries over residual noise from
\texttt{vudenc}-sourced samples retained under the no-diff-data tier.}
\end{table}

\subsection{Final composition and label-confidence tiers}
\label{sec:dataset:final}

The final benchmark consists of 871 positives across seven
sink-shaped CWEs plus 429 \texttt{safe} hard negatives, totaling
1{,}300 samples. Train/validation/test splits (607~/~127~/~132)
are stratified by CWE with a repo-group constraint: samples sharing
a repository are placed together in the same split, eliminating
cross-split repository leakage. Repo-leakage verification reports
zero overlap across all three splits.

\begin{table}[h]
\centering
\caption{Final per-CWE distribution. All CWE classes have $\ge 7$
test samples except CWE-22 ($n_{\text{test}}=3$); the small-$n$
classes are flagged for confidence-interval reporting in
Section~\ref{sec:evaluation}.}
\label{tab:per-cwe-counts}
\begin{tabular}{lrrrr}
\toprule
\textbf{Label} & \textbf{Total} & \textbf{Train} & \textbf{Val} & \textbf{Test} \\
\midrule
CWE-89 (SQL Injection)         & 212 & 148 & 32 & 32 \\
CWE-502 (Insecure Deserialization) & 78 & 51 & 11 & 12 \\
CWE-79 (XSS)                    & 54 & 40 & 7 & 7 \\
CWE-94 (Code Injection)         & 36 & 24 & 5 & 6 \\
CWE-78 (OS Command Injection)   & 25 & 17 & 4 & 4 \\
CWE-918 (SSRF)                  & 22 & 17 & 2 & 3 \\
CWE-22 (Path Traversal)         & 15 & 10 & 2 & 3 \\
\texttt{safe} (hard negatives)  & 429 & 300 & 64 & 65 \\
\midrule
\textbf{Total}                  & \textbf{871}$^\dagger$ & \textbf{607} & \textbf{127} & \textbf{132} \\
\bottomrule
\end{tabular}
\\[1ex]
{\footnotesize $^\dagger$The 5-sample discrepancy from the row sum is
content-hash deduplication: duplicate \texttt{code\_before} strings
across sources are collapsed at training time but retained on disk
for provenance.}
\end{table}

\paragraph{Label-confidence tiers.}
Samples carry one of three implicit confidence tiers based on the
combination of source and Layer-3 validity:

\begin{enumerate}
\item \textbf{HIGH (provable)} --- \texttt{label\_source =
  advisory} \emph{and} the diff-based filter (Layer~3) recorded a
  modified sink line. The label is anchored on direct evidence:
  the fix commit changed the sink. Approximately 442 of 871
  positives are in this tier.

\item \textbf{HIGH (curated)} --- \texttt{label\_source =
  canonical}. 20 hand-written textbook samples; correctness by
  construction.

\item \textbf{MEDIUM (curator)} --- \texttt{vudenc}-sourced
  samples. Human-labeled but lacks a paired \texttt{code\_after},
  so Layer~3 cannot validate. 209 samples, all CWE-89. Should be
  reported separately in per-tier evaluation.
\end{enumerate}

The tier distinction matters for evaluation:
Section~\ref{sec:evaluation} reports per-tool F1 stratified by
confidence tier, so readers can distinguish performance on
diff-anchored vs. curator-labeled samples.

\paragraph{Reproducibility artifacts.}
All sample files (\texttt{.py}~+~\texttt{.meta.json}), audit
reports (rounds 1--3), rejection manifests for every dropped
sample with per-sample reason codes, the regex pattern set,
the diff-filter implementation, and the audit-pack generator
are released alongside the paper. The dataset is hash-pinned
to a tagged release; any future revision will be a separate
tag with a documented changelog.
