\section{Evaluation}
\label{sec:evaluation}

This section applies the benchmark to a set of detectors. We do not
rank detectors here. The purpose of the evaluation is to demonstrate
what the benchmark measures: detection accuracy on sink-validated
labels, and the robustness drop that the seven adversarial mutators
expose.

\subsection{Research questions}
\label{sec:eval:rq}

\begin{description}
  \item[RQ1] How accurately do free static analysis tools classify the
    seven sink-shaped CWE classes on the clean test set?
  \item[RQ2] How does each tool's accuracy change when the test
    samples are rewritten by a semantics-preserving mutator? Which
    mutator, if any, exposes the most brittleness?
  \item[RQ3] Does the diff-based label filter (Section~\ref{sec:dataset})
    change the picture a detector evaluation paints? That is, are the
    tools being scored against labels they could plausibly satisfy?
\end{description}

\subsection{Setup}
\label{sec:eval:setup}

\paragraph{Test set.} The held-out test split is 129 samples: 67
vulnerable positives spanning the seven CWE classes and 62 \texttt{safe}
hard negatives. Three hard negatives from the underlying split are
excluded from active evaluation because their source files exceed the
practical context window of current LLM detectors (the largest is a
1.79~MB bundled artifact equivalent to roughly 448K tokens); the
exclusion is applied uniformly across all detectors so the reported
numbers remain like-for-like. The seven single mutators and the composed
mutator are applied to the 67 positives, producing eight adversarial
variant sets of 67 samples each. The hard negatives are not mutated;
they appear only in the \emph{clean} run, where they measure false
positives.

\paragraph{Metrics.} Every detector prediction is scored one-vs-rest per
CWE class: a sample labelled CWE-X that the tool flags as CWE-X is a
true positive for that class, a sample of any other label flagged as
CWE-X is a false positive, and so on (Section~\ref{sec:methodology}).
Each detector is reported on five bounded statistics rather than a
single composite score. Macro-F1 averages per-class F1 with equal
weight across the seven CWE classes; macro-averaging matters here
because class sizes range from 3 (CWE-22) to 32 (CWE-89). Detection
MCC~\cite{chicco2020mcc} is the Matthews correlation coefficient on the
binary vulnerable-versus-safe collapse of the same predictions; it lies
in $[-1, +1]$, equals zero at chance agreement, and is the recommended
imbalance-robust replacement for accuracy on this kind of corpus.
Severity-Weighted Recall (SWR) is the CVSS-weighted fraction of
positives the detector flagged, using each advisory's published CVSS
base score as the weight; positives without a CVSS score are excluded
from the SWR pool and the pool size is reported alongside the
statistic. Hierarchical CWE Macro-Accuracy (HCMA) replaces flat
exact-match accuracy with Wu-Palmer~\cite{wupalmer1994} similarity on
a hand-coded subtree of MITRE CWE 4.14, so that close-sibling
mis-classifications (e.g.\ predicting CWE-94 when the truth is
CWE-78, both under CWE-74 Injection) receive partial credit. Weighted
Cohen's $\kappa$~\cite{cohen1968kappa} aggregates the same partial
credit into a single chance-corrected agreement number with
disagreement weights $1 - \text{Wu-Palmer}$ between two CWEs and $1$
between \emph{safe} and any CWE. The five measures answer separate
questions (CWE identification, binary detection, severity-weighted
recall, hierarchical-near-miss credit, and chance-corrected
agreement); they are reported jointly rather than combined into a
single composite, following the rationale of the project's evaluation
methodology document.

\paragraph{Robustness drop.} For a tool and a mutator, the robustness
drop is the macro-F1 on the clean positives minus the macro-F1 on that
mutator's output. Because the hard negatives are absent from the
mutator variants, the clean baseline for this comparison is recomputed
over the 67 positives only, so the subtraction is like-for-like. Per
mutator family (sink-preserving, sink-targeted, dataflow) we report
both the per-mutator drop and, for sink-preserving, the mean across
its four mutators.

\paragraph{Detectors.} We report seven detectors spanning four tiers.
Two free, locally-installable pattern-matching SAST tools: Bandit~1.9.4
and Semgrep~1.163.0. One cloud-backed, ML-augmented SAST tool: Snyk
Code~1.1305.0, included to test the snippet-format fit of a
whole-project dataflow detector against the on-disk function extracts.
Three frontier LLM detectors: Claude Sonnet 4.6 (Anthropic), GPT-4o
(OpenAI), and DeepSeek R1 (via OpenRouter, paid endpoint). One
benchmark-trained learned detector: a fine-tuned GraphCodeBERT
classifier (training schedule in Section~\ref{sec:eval:model}). Semgrep
is run with the \texttt{p/python}, \texttt{p/security-audit}, and
\texttt{p/owasp-top-ten} registry rule packs. Each SAST tool emits
findings that we map to the seven CWE classes; for Bandit the mapping
is by rule identifier rather than by Bandit's own \texttt{issue\_cwe}
field, which mislabels \texttt{eval} as CWE-78 and \texttt{yaml.load}
as CWE-20; for Snyk Code and Semgrep the mapping is read from each
rule's CWE tags. LLM detectors share a fixed classification prompt;
each response's first non-empty line is parsed as the prediction, with
a reasoning-model fallback that scans backward for the verdict line if
the model leads with chain-of-thought prose.

\subsection{Headline results}
\label{sec:eval:headline}

Table~\ref{tab:pyvulsev-results} summarizes detector performance on the
PyVulSev benchmark. Each detector is reported on macro-F1 over the
129-sample clean test set, detection MCC on the binary
vulnerable-versus-safe collapse of the same predictions, and
Severity-Weighted Recall on the 27-sample CVSS-scored pool of clean
positives. The Comp.\ column gives macro-F1 on the composed-mutator
variant of the 67 positives. The $\Delta$F1$^{\star}$ column reports
the worst-case (largest) robustness drop in macro-F1 that any single
mutator induces, computed on the positives only so that the
subtraction is like-for-like; a value near zero means the detector is
unaffected by any of the seven mutators in isolation, and a large
positive value identifies the most brittle axis for that detector. The full per-family breakdown is in
Table~\ref{tab:robustness-drop}.

\begin{table*}[t]
\centering
\caption{PyVulSev results on the 129-sample clean test set.
Tool versions: Bandit~1.9.4, Semgrep~1.163.0, Snyk Code~1.1305.0,
GraphCodeBERT \texttt{phase3} checkpoint, Claude Sonnet 4.6, GPT-4o,
DeepSeek R1 (OpenRouter \texttt{deepseek/deepseek-r1}, paid endpoint). Macro-F1 is averaged across the seven
CWE classes under one-vs-rest scoring. MCC is the Matthews
correlation coefficient on the binary vulnerable-versus-safe
collapse. SWR is Severity-Weighted Recall on the 27 clean positives
that carry a CVSS base score. HCMA is the hierarchical CWE
macro-accuracy under Wu-Palmer~\cite{wupalmer1994} similarity on
the hand-coded subtree of MITRE CWE 4.14
(\texttt{src/eval/cwe\_map.py}). $\kappa_w$ is the weighted Cohen's
$\kappa$ over the eight-label space (seven CWEs plus \emph{safe})
with disagreement weights $1{-}$Wu-Palmer between two CWEs and $1$
between \emph{safe} and any CWE. Comp.\ is macro-F1 on the
composed-mutator variant of the 67 positives.
$\Delta$F1$^{\star}$ is the worst-case drop in macro-F1
across the seven single mutators (Table~\ref{tab:robustness-drop}),
computed on the positives only.}
\label{tab:pyvulsev-results}
\small
\setlength{\tabcolsep}{4.5pt}
\begin{tabular}{l r r r r r r r}
\hline
Name & macro-F1 & MCC & SWR & HCMA & $\kappa_w$ & Comp. & $\Delta$F1$^{\star}$ \\
\hline
Bandit               & 0.510 & +0.478 & 0.745 & 0.564 & +0.567 & 0.589 & +0.187 \\
Semgrep              & 0.315 & +0.157 & 0.634 & 0.353 & +0.227 & 0.375 & +0.132 \\
Snyk Code            & 0.063 & $-$0.235 & 0.143 & 0.069 & $-$0.097 & 0.088 & +0.036 \\
\hline
Claude Sonnet 4.6    & 0.585 & +0.117 & 0.903 & 0.805 & +0.396 & 0.867 & $0.000$ \\
GPT-4o               & 0.356 & +0.047 & 0.420 & 0.419 & +0.295 & 0.565 & +0.026 \\
DeepSeek R1          & 0.596 & +0.410 & 0.691 & 0.646 & +0.539 & 0.567 & +0.068 \\
\hline
GraphCodeBERT (ours) & 0.524 & +0.556 & 0.870 & 0.547 & +0.548 & 0.603 & $0.000$ \\
\hline
\end{tabular}
\end{table*}

The seven detectors separate on different axes; no single tool leads
on every measure. DeepSeek R1 narrowly leads on macro-F1 (0.596,
ahead of Sonnet 4.6 at 0.585 and the trained GraphCodeBERT at 0.524).
Sonnet 4.6 leads on severity-weighted recall (0.903 — 90\% of
CVSS-weighted danger on the 27-sample pool) and on the hierarchical
HCMA measure (0.805). GraphCodeBERT leads on detection MCC
($+0.556$), the binary-collapse statistic; DeepSeek R1 is the only
LLM with a comparable MCC ($+0.410$), while Sonnet ($+0.117$) and
GPT-4o ($+0.047$) sit far below the SAST and learned tiers on this
axis because both LLMs over-flag hard negatives. Bandit edges all
other detectors on weighted Cohen's $\kappa$ ($+0.567$); Bandit and
Semgrep occupy distinct points on the precision-recall plane:
Bandit combines balanced precision and recall on the classes it
covers, while Semgrep's stricter taint-shape rules drive precision
up but miss many fragmentary positives, particularly on CWE-89.
Snyk Code is an outlier on every detection metric — macro-F1 0.063,
MCC $-0.235$, SWR 0.143 — and we treat the row as a methodological
signal rather than a tool ranking; the discussion in
Section~\ref{sec:eval:llm} returns to why a snippet-format benchmark
is unfair to whole-project dataflow detectors like Snyk Code and
CodeQL (see Section~\ref{sec:threats:external}).

The hierarchical and agreement columns sharpen the comparison
between Bandit and the trained detector. On macro-F1 Bandit
(0.510) trails GraphCodeBERT (0.524) by 0.014, but on HCMA the gap
inverts and Bandit (0.564) edges ahead of GraphCodeBERT (0.547)
by 0.017. The mechanism is visible in the per-CWE breakdown:
Bandit's confusions are dominated by close-sibling errors within
the injection subtree (CWE-94 samples flagged as CWE-78 via the
\texttt{B602}/\texttt{B603} subprocess rules; CWE-78 samples
flagged as CWE-94 via the \texttt{B102}/\texttt{B307}
\texttt{exec}/\texttt{eval} rules), which Wu-Palmer credits as
partial matches but macro-F1 scores as flat misses. GraphCodeBERT's
confusions are more diffuse and benefit less from partial credit.
Weighted $\kappa$ converges on the same conclusion through a
different route: Bandit and GraphCodeBERT are within 0.019 of each
other ($+0.567$ vs $+0.548$), and both lead Semgrep by more than
0.30. The single-number reading is that Bandit and GraphCodeBERT
are comparable detectors on clean code under any partial-credit
metric — the trained detector's headline win is most visible on
SWR and disappears on the chance-corrected agreement statistic.

The robustness column tells a sharper story. The two SAST tools
each have a worst-case drop above 0.13 macro-F1: Bandit loses 0.187
under \texttt{sink\_attr\_obfuscate} and Semgrep loses 0.132 under
\texttt{sink\_via\_globals}. The trained detector's worst-case drop
is $0.000$ — it is invariant to every mutator in the suite.
Section~\ref{sec:eval:robustness} unpacks this contrast by mutator
family: the SAST cliffs come from the sink-targeted family, all
three detectors are invariant to the four sink-preserving mutators,
and the dataflow mutator is invariant on every detector reported
here because none of the three performs dataflow analysis.

The three LLM rows tell a different story from the SAST and learned
detectors, and they separate sharply from each other. DeepSeek R1
takes the macro-F1 lead (0.596) and pairs it with a relatively high
MCC ($+0.410$), indicating that its CWE assignments and its
binary vulnerable/safe decision agree more often than either Sonnet's
or GPT-4o's. Sonnet 4.6 ranks just below DeepSeek on macro-F1 (0.585)
but takes the SWR and HCMA leads of the entire table; its MCC
($+0.117$) is low because it over-flags hard negatives as CWE-22 and
CWE-918. GPT-4o sits at the opposite extreme: macro-F1 (0.356) below
Bandit's, MCC ($+0.047$) near chance, SWR (0.420) the lowest of the
six well-fit detectors (excluding Snyk Code; see
Section~\ref{sec:eval:wholeproject}). Section~\ref{sec:eval:llm} unpacks all three
findings by CWE. The robustness columns are flat for every LLM:
Sonnet's worst-case drop is $0.000$, GPT-4o's is $+0.026$, DeepSeek
R1's is $+0.068$. Reporting clean and post-mutation scores side by
side extends the CASTLE layout, which reports detection only, with
the robustness axis the seven mutators measure.

\subsection{Detection on the clean test set (RQ1)}
\label{sec:eval:clean}

Table~\ref{tab:clean-percwe} gives the per-class F1 of the two
pattern-matching SAST tools and the trained GraphCodeBERT detector
on the clean test set. The three LLM detectors and Snyk Code are
discussed by CWE in their respective subsections
(Sections~\ref{sec:eval:llm} and~\ref{sec:eval:wholeproject}).

\begin{table}[h]
\centering
\caption{Per-CWE F1 on the clean test set. $n$ is the number of
positive test samples in the class.}
\label{tab:clean-percwe}
\begin{tabular}{l r r r r}
\hline
CWE & $n$ & Bandit & Semgrep & GCB \\
\hline
CWE-22  Path Traversal        &  3 & 0.00 & 0.00 & 0.00 \\
CWE-78  OS Command Injection  &  4 & 0.38 & 0.29 & 0.40 \\
CWE-79  Cross-site Scripting  &  7 & 0.60 & 0.33 & 0.67 \\
CWE-89  SQL Injection         & 32 & 0.80 & 0.06 & 0.63 \\
CWE-94  Code Injection        &  6 & 0.92 & 0.80 & 0.57 \\
CWE-502 Insecure Deserial.    & 12 & 0.83 & 0.73 & 0.83 \\
CWE-918 SSRF                  &  3 & 0.00 & 0.00 & 0.57 \\
\hline
Macro-F1 (positives + safe)   &    & 0.51 & 0.32 & 0.52 \\
\hline
\end{tabular}
\end{table}

CWE-22 (path traversal) scores zero across all three detectors:
neither Bandit nor Semgrep ships a usable path-traversal rule for the
test samples, and the trained detector has only three CWE-22 positives
in the test set, too few to learn from given the 15 CWE-22 samples in
the entire training corpus. CWE-918 (SSRF) separates the detectors:
both SAST tools score zero, but the trained GraphCodeBERT detector
recovers F1 0.57. Bandit's closest CWE-918 signal, the
\texttt{urllib.urlopen} check, produced five false positives and no
true positives. The benchmark therefore covers one CWE class
(CWE-918) on which free static analysis fails and a learned detector
trained on the benchmark does not, and the macro-average reports the
remaining CWE-22 gap rather than averaging it away.

The detectors diverge most on SQL injection. Bandit reaches F1 0.80
on CWE-89; the trained detector reaches 0.63; Semgrep reaches 0.06,
flagging one of 32 positives. The cause is the composition of the
CWE-89 sub-corpus. Most of these samples are fragments drawn from
the \texttt{vudenc} source and lack the framework-recognisable
source-to-sink shape that Semgrep's taint rules require. Bandit's
\texttt{B608} rule matches inline SQL string construction lexically
and fires readily on such fragments. The trained detector, fine-tuned
on the same fragments, sits between the two: it identifies fragmentary
SQL injection more reliably than Semgrep's taint-shape rules but
trails Bandit's broader lexical net. The divergence isolates lexical
matching, taint-style reasoning, and learned representation on
exactly the inputs where the three approaches disagree.

The headline macro-F1 of 0.51 (Bandit), 0.32 (Semgrep) and 0.52
(trained) is computed over all 129 clean samples, including the 62
hard negatives. Restricted to the 67 positives the same detectors
score 0.59, 0.38 and 0.57; the gap between the all-129 and
positives-only numbers is the F1 cost of false positives raised on
safe code. Bandit's gap is the largest at 0.08 F1 points, chiefly
reflecting low-severity \texttt{subprocess} warnings (\texttt{B603},
\texttt{B607}) firing on benign command-execution code; the trained
detector's gap is the smallest at 0.05 points, indicating a lower
false-positive rate on hard negatives than either SAST tool achieves.

\subsection{Adversarial robustness (RQ2)}
\label{sec:eval:robustness}

The seven single mutators fall into three families. The
\emph{sink-preserving} family — \texttt{dead\_code\_injection} (DC),
\texttt{string\_split} (SS), \texttt{variable\_rename} (VR),
\texttt{wrapper\_extraction} (WE) — modifies code around the sink but
leaves the sink identifier (\texttt{os.system},
\texttt{cursor.execute}, \texttt{eval},~\ldots) intact and the
argument expressions reaching the sink unchanged. The
\emph{sink-targeted} family — \texttt{sink\_attr\_obfuscate} (SAO),
\texttt{sink\_via\_globals} (SVG) — rewrites the sink call itself.
SAO replaces attribute-style sinks \texttt{os.system(x)} with
\texttt{getattr(os, "system")(x)} (CWE-78/89/918/502), and SVG
replaces bare-builtin sinks \texttt{eval(x)} with
\texttt{\_\_import\_\_("builtins").\_\_dict\_\_["eval"](x)}
(CWE-94). The \emph{dataflow} family —
\texttt{taint\_through\_dict} (TTD) — leaves both the surrounding
code and the sink identifier intact, and rewrites the first
positional argument of every sink call as a one-entry
dict-construction/subscript pair, breaking any direct static
lineage from a parameter to the sink argument. All three
sink-targeted and dataflow mutators are held out of training-time
augmentation alongside \texttt{string\_split}.
Table~\ref{tab:robustness} reports macro-F1 across all seven single
variants and the composed variant; Table~\ref{tab:robustness-drop}
reports the robustness drops grouped by family.

\begin{table}[h]
\centering
\caption{Macro-F1 by variant, computed over the 67 positives.
DC: dead code, SS: string split, VR: variable rename, WE: wrapper
extraction, SAO: sink attribute obfuscation, SVG: sink via globals,
TTD: taint through dict, Comp: composed. Sink-preserving mutators
are columns DC--WE; sink-targeted are SAO and SVG; dataflow is TTD.}
\label{tab:robustness}
\small
\setlength{\tabcolsep}{3.5pt}
\begin{tabular}{l r r r r r r r r r}
\hline
Tool & Clean & DC & SS & VR & WE & SAO & SVG & TTD & Comp \\
\hline
Bandit         & 0.59 & 0.59 & 0.59 & 0.59 & 0.59 & 0.40 & 0.45 & 0.59 & 0.58 \\
Semgrep        & 0.38 & 0.38 & 0.38 & 0.38 & 0.38 & 0.28 & 0.25 & 0.37 & 0.38 \\
GraphCodeBERT  & 0.57 & 0.59 & 0.60 & 0.60 & 0.61 & 0.60 & 0.57 & 0.60 & 0.60 \\
\hline
\end{tabular}
\end{table}

\begin{table}[h]
\centering
\caption{Robustness drop (clean $-$ mutator) by family.
$\Delta$F1\textsubscript{sink-preserve} is the mean of the four
sink-preserving mutators; $\Delta$F1\textsubscript{SAO},
$\Delta$F1\textsubscript{SVG}, and $\Delta$F1\textsubscript{TTD}
are reported individually because each covers a different sink
position or flow path. $\Delta$SWR is the corresponding drop in
severity-weighted recall on the 27-sample CVSS pool.}
\label{tab:robustness-drop}
\small
\setlength{\tabcolsep}{3.5pt}
\begin{tabular}{l r r r r r r r r}
\hline
& \multicolumn{2}{c}{Sink-preserve} & \multicolumn{2}{c}{SAO} & \multicolumn{2}{c}{SVG} & \multicolumn{2}{c}{TTD} \\
Tool & $\Delta$F1 & $\Delta$SWR & $\Delta$F1 & $\Delta$SWR & $\Delta$F1 & $\Delta$SWR & $\Delta$F1 & $\Delta$SWR \\
\hline
Bandit         & $+$0.001 & $0.000$  & $+$0.187 & $+$0.347 & $+$0.143 & $+$0.246 & $0.000$  & $0.000$  \\
Semgrep        & $+$0.005 & $0.000$  & $+$0.109 & $+$0.196 & $+$0.132 & $+$0.246 & $+$0.018 & $-$0.029 \\
GraphCodeBERT  & $-$0.036 & $+$0.009 & $-$0.035 & $+$0.034 & $0.000$  & $0.000$  & $-$0.030 & $0.000$  \\
\hline
\end{tabular}
\end{table}

The three detectors separate sharply on the sink-targeted mutators.
The pattern-matching SAST tools both show a robustness cliff: Bandit
drops 0.187 macro-F1 under SAO and 0.143 under SVG, with
severity-weighted recall drops of 0.347 and 0.246 — the SAO drop is
a 47\% relative loss in SWR. Semgrep drops 0.109 and 0.132 macro-F1,
and 0.196 and 0.246 SWR. The drops are consistent across the two
sink-targeted mutators, evidence that the cliff is a property of the
sink-token axis rather than an artifact of a single rewriting trick.
The trained GraphCodeBERT detector is invariant to both sink-targeted
mutators ($\Delta$F1 $-$0.035 and $0.000$; $\Delta$SWR $+$0.034 and
$0.000$) and to the four sink-preserving mutators. Within the
per-sample F1 swing of approximately $1.5\%$ that the 67-positive
test set produces, the learned model treats every variant of a
sample as the same class.

On the four sink-preserving mutators all three detectors are
invariant (mean $\Delta$F1 between $-$0.036 and $+$0.005, no single
mutator larger than 0.05). The held-out \texttt{string\_split}
produces a change indistinguishable from the three training-augmented
mutators on the learned detector, supporting the conclusion that
these four mutators occupy a common transformation axis: each
preserves the sink token, so the lexical SAST tools see an unchanged
sink and the learned detector, trained on similar augmentations,
sees a distribution it has already adapted to.

The dataflow mutator (TTD) is invariant on all three detectors
($\Delta$F1 between $-$0.030 and $+$0.018; $\Delta$SWR between
$-$0.029 and $0.000$). We read this as a property of the current
detector set rather than a property of the TTD probe: Bandit
performs no dataflow analysis, the Semgrep configurations used here
do not enable taint mode, and the trained detector was not
augmented with dataflow-perturbing variants during fine-tuning, so
all three see the dict-wrapped argument as effectively equivalent
to the original. The TTD column is included in the benchmark
because the third axis is real even when the current detectors do
not exercise it; it provides a slot for future dataflow-aware
detectors (CodeQL, taint-mode Semgrep configurations, a learned
detector trained on dict-flow augmentations) where the cliff would
appear if the detector relies on lineage-tracking that does not
propagate through dict reads.

The contrast between the three families is the central robustness
finding of this work. Mutator design must target sink syntax to
exercise robustness against pattern-matching SAST; semantic features
beyond the sink identifier are what carry the learned detector
through that axis; and the dataflow axis is the right place to look
for a robustness cliff in a third generation of detectors that this
work does not evaluate.

The SVG family is currently applicable to a small portion of the
test set: only 6 of the 67 positives use a bare builtin sink (all
CWE-94, all using \texttt{eval}); SAO applies to 46 and TTD to 49.
The SVG drop numbers are therefore noisier than the SAO numbers,
and we report them because they corroborate the direction and
magnitude of the SAO cliff on the same two SAST tools and the same
flat response on the learned detector. Section~\ref{sec:threats}
discusses the implications for benchmark coverage and the remaining
sink-targeted axis (dynamic dispatch on receiver objects) that this
benchmark does not yet exercise.

\subsection{Whole-project dataflow detectors}
\label{sec:eval:wholeproject}

Snyk Code is a cloud-backed, ML-augmented SAST product designed for
whole-application analysis: the engine ingests an entire repository,
resolves cross-file dataflow, and reports findings keyed to recognised
framework sources. On the function-extract benchmark we report here it
underperforms every other detector on every detection metric (macro-F1
0.063, MCC $-0.235$, SWR 0.143, HCMA 0.069). Two observations together
explain the result. First, the per-CWE breakdown shows that Snyk Code
flags 1 of 32 CWE-89 positives, 0 of 12 CWE-502 positives, and 0 of 4
CWE-78 positives — the classes whose flow shapes most depend on the
framework-source-to-application-sink lineage that whole-project
analysis is built to track. Second, the detector produces a high
volume of false positives on the hard negatives (11 false positives on
CWE-22, 11 on CWE-79, 5 on CWE-918), consistent with a dataflow engine
guessing at sources when none are recognised. Like the LLM and learned
detectors, Snyk Code is essentially invariant to the seven mutators
(worst-case macro-F1 drop $+0.036$ under SVG), so the result is not a
robustness story; it is a shape-fit story. We attempted CodeQL~2.25.5
with the \texttt{codeql/python-queries} suite as a second whole-project
tool and observed the same root cause in its strongest form: CodeQL
returned zero findings on the clean set, because its source model
requires application-level entry points (Flask routes, Django views)
that are absent from the function extracts. We document the CodeQL
attempt and the shape-fit conclusion in
Section~\ref{sec:threats:external} as an external-validity bound:
single-file extracts are not a fair benchmark format for
dataflow-aware detectors whose semantics require whole-application
context. The Snyk Code row appears in Table~\ref{tab:pyvulsev-results}
to make this bound empirically visible rather than only argumentative.

\subsection{LLM detectors}
\label{sec:eval:llm}

The harness evaluates LLM detectors through the same interface: a
fixed system prompt lists the seven CWE classes, the file contents
form the user turn, and the first non-empty line of the response is
parsed as the prediction. A full pass over the 129 clean samples and
the eight mutator variants (seven single mutators plus composed) is
665 model calls per detector. We evaluate three LLMs: Claude
Sonnet 4.6 (Anthropic), GPT-4o (OpenAI), and DeepSeek R1 (via
OpenRouter, paid endpoint). All three share the same system prompt
and the same one-line-CWE parsing rule; the implementation is one
Messages-API or Chat-Completions call per sample, with
provider-specific rate-limit handling and a per-sample error fallback
that records an empty prediction rather than crashing the sweep.
DeepSeek R1 is a reasoning model that emits a chain-of-thought
before the answer; the parser strips inline \texttt{<think>...</think>}
blocks (OpenRouter usually places them in a separate
\texttt{message.reasoning} field, but not always) and reads the first
verdict-shaped line.

\paragraph{Claude Sonnet 4.6.} On macro-F1, Sonnet leads the table at
0.585 (against 0.524 for GraphCodeBERT, 0.510 for Bandit, 0.356 for
GPT-4o, 0.315 for Semgrep). Severity-Weighted Recall is also the
highest in the table at 0.903 on the 27-sample CVSS pool — Sonnet
catches the largest fraction of CVSS-weighted danger of any detector
reported. Detection MCC, however, is only $+0.117$, the second-lowest
in the table. The mechanism is visible in the per-CWE breakdown:
Sonnet over-flags hard negatives as CWE-22 (15 false positives against
the 3 CWE-22 positives) and CWE-918 (17 false positives against 3
positives), so the binary vulnerable-versus-safe collapse that
underpins MCC absorbs a heavy false-positive rate that macro-F1
averages across the seven CWE classes. Hierarchical accuracy (HCMA
0.805) and weighted $\kappa$ ($+0.396$) are both the highest LLM
values in the table; Sonnet's confusions stay close to the correct
CWE subtree even when it over-flags.

\paragraph{GPT-4o.} GPT-4o is the weakest of the three LLM detectors
on every clean-set metric: macro-F1 0.356 (below Bandit's 0.510), MCC
$+0.047$ (near chance), HCMA 0.419, and SWR 0.420. The aggregate
result masks a sharp per-class imbalance. GPT-4o reaches F1 0.915 on
CWE-89 — higher than every other detector in the table except DeepSeek
R1's 0.935 — but scores zero on CWE-94, missing all six code-injection
positives, and produces 15 and 16 false positives on CWE-22 and
CWE-918 respectively. The pattern suggests a miscalibration rather
than a systemic weakness: GPT-4o recognizes the canonical SQL-injection
shape but over-applies path-traversal and SSRF labels to safe code
and fails to apply the code-injection label at all.

\paragraph{DeepSeek R1.} DeepSeek R1 posts the highest macro-F1 in
the table (0.596, narrowly above Sonnet) and the highest CWE-89 F1
of any detector (0.935 against 32 positives, exceeding GPT-4o's 0.915
and the trained GraphCodeBERT's 0.625). Its detection MCC ($+0.410$)
is by far the highest of the three LLMs and approaches the SAST and
learned tiers (Bandit $+0.478$, GraphCodeBERT $+0.556$), indicating
that its CWE assignments and its binary vulnerable/safe decision are
more consistent than the other LLMs'. The per-CWE breakdown shows
strong performance on the four high-support classes (CWE-89 F1 0.935,
CWE-94 0.923, CWE-78 0.857, CWE-502 0.700) with the same
over-flagging pattern on CWE-22 (10 false positives, F1 0.308) and
CWE-918 (10 false positives, F1 0.143) that the other LLMs exhibit
on those small classes. Treating CWE-22 and CWE-918 as conditional
labels rather than absolute classifications would close most of the
calibration gap between DeepSeek R1 and the trained detector, but
that re-scoring is not part of the harness reported here.

\paragraph{Robustness.} All three LLMs are essentially invariant to
the seven mutators. Sonnet's worst-case macro-F1 drop is exactly
$0.000$, GPT-4o's is $+$0.026 under SAO, and DeepSeek R1's is
$+$0.068 under \texttt{string\_split} — the largest LLM drop in the
table but still well below the SAST cliffs ($+$0.187 for Bandit
under SAO, $+$0.132 for Semgrep under SVG). Like the trained
GraphCodeBERT detector, all three LLMs treat each mutator variant
of a sample as the same class as the original. The sink-token cliff
that pattern-matching SAST tools exhibit
(Section~\ref{sec:eval:robustness}) is absent for every learned or
LLM detector reported here; the cliff is a property of lexical
sink-matching rather than a property of the mutator suite. The
caveats from the learned-detector discussion apply equally to the
LLM rows: SVG applies to only 6 of 67 positives, and the dataflow
probe (TTD) is invariant on every non-SAST detector evaluated.

\subsection{Trained-model detector}
\label{sec:eval:model}

The benchmark's training split (607 samples) is used to fine-tune a
dual-task GraphCodeBERT~\cite{graphcodebert} classifier that predicts
the CWE class and the CVSS base score from \texttt{code\_before}.
Training uses a three-phase schedule: a three-epoch supervised
contrastive warmup, five epochs of LDAM~\cite{cao2019ldam} margin
loss on a class-balanced sampler, and two epochs of deferred
re-weighting~\cite{cao2019ldam} with Effective-Number~\cite{cui2019cb}
class weights. Online augmentation applies three of the seven
benchmark mutators (\texttt{dead\_code\_injection},
\texttt{variable\_rename}, \texttt{wrapper\_extraction}) during
training; the other four (\texttt{string\_split},
\texttt{sink\_attr\_obfuscate}, \texttt{sink\_via\_globals},
\texttt{taint\_through\_dict}) are held out as test-time robustness
probes — \texttt{string\_split} for the sink-preserving axis,
\texttt{sink\_attr\_obfuscate} and \texttt{sink\_via\_globals} for
the sink-token axis, and \texttt{taint\_through\_dict} for the
dataflow axis. The best checkpoint is selected by validation
macro-F1.

The trained detector reaches macro-F1 0.524 and detection MCC
$+0.556$ on the clean test set, leading both pattern-matching SAST
baselines (Bandit MCC $+0.478$, Semgrep $+0.157$) and all three LLM
detectors (DeepSeek R1 $+0.410$, Sonnet $+0.117$, GPT-4o $+0.047$)
on the chance-corrected binary statistic. The advantage on MCC
reflects the trained detector's lower false-positive rate on hard
negatives — DeepSeek R1 (macro-F1 0.596) and Sonnet (macro-F1 0.585)
both exceed the trained detector's 0.524 on the per-CWE macro
average, but both LLMs over-flag \texttt{safe} samples as CWE-22 and
CWE-918, so the binary collapse rewards GraphCodeBERT's calibration
even where it loses on the per-CWE macro. Per-CWE coverage is also distinctive: the trained detector
recovers a non-zero CWE-918 signal (F1 0.57) on which the two SAST
tools produce no true positives, and trails Bandit on CWE-89 and
CWE-94 (Table~\ref{tab:clean-percwe}). Severity-weighted recall is
0.870 on the 27-sample CVSS pool — second only to Sonnet's 0.903,
and ahead of Bandit (0.745), Semgrep (0.634), and GPT-4o (0.420).
We treat this as a defensible small-data result rather than a
statement about a state-of-the-art detector: 607 training samples
is small for a 126.6M-parameter model, and reviewers should read
the result as an existence proof that a benchmark-tuned learned
detector can compete with mature SAST baselines and frontier LLMs
on Python sink-shaped CWEs, not as evidence that the trained
detector generalizes outside this benchmark.

Under the four sink-preserving mutators the trained detector's
macro-F1 changes by between $-$0.024 and $-$0.044 (mean $-$0.036),
i.e. the detector improves slightly under mutation; the held-out
\texttt{string\_split} produces a $-$0.037 change, similar in
magnitude to the three augmented mutators. Under the two sink-targeted
mutators the change is $-$0.035 (SAO) and $0.000$ (SVG) in macro-F1
and $+$0.034 and $0.000$ in SWR. Under the dataflow mutator (TTD)
the change is $-$0.030 in macro-F1 and $0.000$ in SWR. The learned
detector is therefore flat across all three mutator families, in
contrast to the pattern-matching SAST tools which drop
0.10--0.19 F1 and 0.20--0.35 SWR on the sink-targeted family
(Section~\ref{sec:eval:robustness}). We read this as evidence that
the learned detector relies on semantic features beyond the sink
identifier and beyond a direct parameter-to-sink lineage, but
caution that the SVG family applies to only 6 of 67 positives and
that the dataflow probe is invariant on every detector reported
here. Section~\ref{sec:threats} treats both points as scope
caveats on the per-family numbers.

\subsection{Per-framework breakdown}
\label{sec:eval:framework}

Every benchmark sample carries a \texttt{framework} field
(Section~\ref{sec:dataset:schema}) identifying which Python web
framework the source file uses: Flask, Django, FastAPI, Starlette,
aiohttp, Tornado, Bottle, or none for non-web code. The released
prediction artifacts (\texttt{reports/eval/*\_predictions.jsonl})
carry one record per (detector, variant, sample) and can be
re-aggregated by framework off-line; we omit the per-framework table
from this presentation because the SAST detectors apply largely
framework-agnostic regex patterns and the LLM and learned-detector
per-framework counts fall within the per-class confidence intervals
already flagged in Section~\ref{sec:threats}. The artifact-level
breakdown remains available in the released JSONL files for
downstream analysis.

\subsection{Effect of the label filter (RQ3)}
\label{sec:eval:labels}

The numbers above are computed against the diff-filtered labels of
Section~\ref{sec:dataset}, whose audited false-positive rate is 26\%.
Run against the unfiltered labels, every tool's apparent precision
falls by construction: a tool that correctly declines to flag a
mislabelled sample is charged a false negative. The filter therefore
does not flatter the tools. It removes samples on which a correct
detector and the label disagree because the label is wrong. Detector
accuracy against sink-validated labels separates this benchmark from
an evaluation against raw CVE-derived tags, where 30 to 50\% of the
ground truth is noise and a tool is partly scored on its agreement
with that noise.
