\section{Introduction}
\label{sec:introduction}

The bulk of automated vulnerability-detection research over the past
five years has focused on C and C++. Recent benchmark and dataset
efforts including BigVul~\cite{bigvul},
CVEfixes~\cite{cvefixes}, CrossVul~\cite{crossvul},
DiverseVul~\cite{diversevul}, and CASTLE~\cite{castle}
concentrate on memory-safety bugs, buffer overflows, and
use-after-free patterns that dominate C-language CVE feeds. Python
receives substantially less attention despite being the dominant
language for security-critical infrastructure: machine-learning
pipelines, scientific computing, web back-ends, automation tooling,
and cloud-native applications. The vulnerability classes that matter
for Python software are sink-anchored in ways the C-centric literature
does not address: SQL injection through ORM and DB-API patterns,
deserialization of pickle and YAML data, code injection via
\texttt{eval} and \texttt{exec}, and server-side request forgery in
webhook endpoints.

When researchers do build Python vulnerability datasets, they
typically inherit labels directly from CVE fix commits. This strategy
is known to produce 25 to 50\% noise from co-changed
files~\cite{croft}. The labeled ``vulnerable'' file in a CVE fix
is not always the actual sink site: commit boundaries cut across many
files, and CVE descriptions rarely identify the precise line where the
vulnerability lives. Recent audits of BigVul~\cite{croft}
quantified this at $\approx 25\%$; informal audits of CVE-fix-derived
data we conducted as part of this work found 52\% mislabeled positives
in the raw, unfiltered CVE-feed corpus. Training and evaluation on
this substrate produces models and tool comparisons that partially
measure noise rather than detection capability.

The CASTLE benchmark~\cite{castle} addressed this evaluation
problem for C/C++ with a hand-curated 250-program corpus and a
comparative study of fourteen static analyzers and ten large language
models. CASTLE accepts CVE/advisory labels at face value but
compensates with small, carefully selected samples. For Python, no
comparable benchmark exists. Researchers fall back to either
CVE-fix-derived datasets with known noise, or
Juliet-style~\cite{juliet} synthetic samples that lack
realistic framework idiom and surrounding application context. Both
substrates limit what can be concluded about tool performance under
real-world conditions. CASTLE measures detection rates only and does
not evaluate robustness against semantics-preserving code
transformations, so its reported scores are an upper bound that may
not survive trivial rewriting of the input.

This paper introduces a Python vulnerability-detection benchmark
covering seven sink-shaped MITRE Top-25~\cite{mitre_top25} CWE classes
(CWE-89, CWE-79, CWE-22, CWE-78, CWE-94, CWE-918, and CWE-502) plus a
``safe'' hard-negative class. We address the label-noise problem with a
three-layer validation methodology that reduces the audited
false-positive rate from 52\% to 26\%, comparable to manually audited
security datasets at a fraction of the human-hour cost. We additionally extend CASTLE's evaluation framing in two
directions: seven semantics-preserving adversarial mutators across
three orthogonal families (sink-preserving, sink-targeted, and
dataflow) isolate specific detection shortcuts, and per-framework
slicing (Flask / Django / FastAPI) exposes framework-specific gaps
in static-analysis rule coverage. The combined benchmark, label-validation methodology,
and adversarial-robustness probe constitute, to our knowledge, the
first Python evaluation substrate with explicit, automated
label-validity criteria.

\paragraph{Contributions}
\begin{itemize}
  \item \textbf{An audit-validated Python vulnerability benchmark.}
    869 samples (440 real-world positives plus 429 synthesized safe
    hard negatives) across seven sink-shaped MITRE
    Top-25~\cite{mitre_top25} CWE classes, with repo-group-stratified
    train / validation / test splits (610 / 127 / 132) and zero
    cross-split repository leakage. Positives are scraped from CVE and
    advisory fix commits; hard negatives are safe twins constructed by
    applying each vulnerability's canonical remediation to its positive,
    preserving surface structure and sink names. Distributed with full
    per-sample metadata, content-hash deduplication, and rejection
    manifests for every dropped sample.

  \item \textbf{A three-layer label-validation methodology.} A pre-ingest
    sink-presence filter, a three-round adversarial audit by independent
    reviewers, and a diff-based filter that requires the sink line to be
    modified between \texttt{code\_before} and \texttt{code\_after}.
    Combined, the three layers reduce audited false-positive rate from
    52\% to 26\%. To our knowledge this is the first CVE-derived Python
    benchmark with explicit, automated label-validity criteria; the
    methodology is reusable across CVE-derived datasets in any language.

  \item \textbf{Seven semantics-preserving adversarial mutators
    across three orthogonal families.} A sink-preserving family
    (dead-code injection, string splitting, identifier rename, wrapper
    extraction) targets shortcuts in code around the sink; a
    sink-targeted family (attribute-access obfuscation,
    builtins-module indirection) rewrites the sink call itself; a
    dataflow mutator routes the tainted argument through an
    intermediate dict subscript. All seven are deterministic per
    sample and AST-level semantics-preserving. By measuring per-tool
    F1 drop per mutator and per family, we surface which shortcut
    each tool depends on, a finer-grained diagnosis than CASTLE's
    aggregate detection rates allow.

  \item \textbf{A comparative evaluation} of seven detectors spanning
    four tiers: pattern-matching SAST (Bandit, Semgrep), cloud-backed
    dataflow SAST (Snyk Code), frontier LLMs (Claude Sonnet 4.6,
    GPT-4o, DeepSeek R1), and a benchmark-trained learned detector
    (a fine-tuned GraphCodeBERT classifier). The harness covers the
    clean test set and the eight adversarial variants (seven single
    mutators plus a composed variant), reporting per-CWE accuracy
    under a five-measure detection profile (macro-F1, detection MCC,
    severity-weighted recall, hierarchical CWE macro-accuracy, and
    weighted Cohen's $\kappa$) and per-mutator robustness drop. The
    headline finding is that the sink-token family separates the
    detector tiers cleanly: Bandit drops 0.187 macro-F1 (a $47\%$
    relative loss in severity-weighted recall) under attribute
    obfuscation and Semgrep 0.132, while the trained and LLM detectors
    are all invariant to every mutator in the suite. The detector
    tiers also separate on clean-set leadership without any single
    tool dominating: DeepSeek R1 narrowly leads macro-F1, Sonnet
    leads severity-weighted recall and hierarchical accuracy,
    GraphCodeBERT leads detection MCC, and Bandit leads weighted
    Cohen's $\kappa$. Snyk Code underperforms every other detector
    on every detection metric, a result we treat as an
    external-validity bound on the benchmark format rather than a
    statement about the tool: single-file function extracts are not
    a fair benchmark for whole-project dataflow detectors.

  \item \textbf{A scope justification} documenting which Top-25 CWEs
    admit sink-anchored detection and which are structural
    (absence-of-code) patterns that require a different methodology
    built around application-level context. We treat the separation
    as a contribution: sink-shaped CWEs admit pattern-based
    evaluation; structural CWEs do not, and benchmarks that mix the
    two confound detector capabilities.
\end{itemize}

The remainder of this paper is organized as follows.
Section~\ref{sec:related} surveys vulnerability datasets, CVE-derived
label noise studies, and the static-analysis / LLM detector landscape.
Section~\ref{sec:dataset} describes dataset construction, sources, the
sink-presence filter, the audit methodology, and the diff-based filter
that together define the benchmark's labels.
Section~\ref{sec:methodology} formalizes the evaluation methodology and
the seven adversarial mutators.
Section~\ref{sec:evaluation} reports the comparative evaluation
results.
Section~\ref{sec:threats} discusses threats to validity, including the
remaining 26\% audit residual.
Section~\ref{sec:conclusion} summarizes findings and outlines future
work, including the deferred structural-CWE evaluation track and
extension to additional languages.
