\section{Adversarial Robustness Methodology}
\label{sec:methodology}

To go beyond ``did the tool detect this vulnerability'' we generate
seven adversarial variants of every test sample. Each variant
rewrites the code in a way that keeps its runtime behavior intact but
breaks a specific shortcut that pattern-matching detectors rely on.
If a tool's F1 score drops sharply on a given mutator's output, that
tool was using the shortcut the mutator targets.

The seven mutators fall into three families. The first four are
\emph{sink-preserving}: \texttt{dead\_code\_injection} (M1),
\texttt{string\_split} (M2), \texttt{variable\_rename} (M3), and
\texttt{wrapper\_extraction} (M4). They alter code around the sink
but leave the sink identifier (\texttt{os.system}, \texttt{eval},
\texttt{cursor.execute},~\ldots) intact and the argument expressions
reaching the sink unchanged. The next two are \emph{sink-targeted}:
\texttt{sink\_attr\_obfuscate} (M5) and \texttt{sink\_via\_globals}
(M6). They rewrite the sink call itself via
\texttt{getattr}-mediated attribute access or \texttt{\_\_dict\_\_}
indexing on the builtins module. The last is a \emph{dataflow}
mutator: \texttt{taint\_through\_dict} (M7) routes the first
positional argument of every sink call through a one-entry dict
literal, preserving the sink identifier but breaking any direct
lineage between a parameter and a sink-argument expression. The
three families probe disjoint detector shortcuts. A detector that
fails on the sink-preserving family is doing positional or
substring reasoning in the code around the sink; a detector that
fails on the sink-targeted family is matching the sink token
itself; a detector that fails on the dataflow family is doing
taint-flow tracking that does not generalise past dict reads.

This section describes the seven mutators, the AST-level invariants
that keep them semantics-preserving, and the way they are composed at
evaluation time. The mutators live in
\texttt{src/red\_team/mutators/} and are deterministic given a sample
identifier and a random seed, so the resulting evaluation is fully
reproducible. Figure~\ref{fig:mutator-tax} lays out the seven mutators
under the three families.

\begin{figure}[h]
\centering
\begin{tikzpicture}[
  font=\footnotesize,
  root/.style={draw, fill=gray!18, align=center, text width=15mm, minimum height=8mm},
  fam/.style={draw, rounded corners, fill=gray!10, align=center, text width=16mm, minimum height=6mm},
  leaf/.style={draw, align=left, text width=44mm, minimum height=5mm, inner sep=3pt},
  arr/.style={-{Stealth[length=1.8mm]}, gray!60},
]
\node[leaf] (m1) {M1\ \ dead-code injection};
\node[leaf, below=1mm of m1] (m2) {M2\ \ string-literal split$^{\dagger}$};
\node[leaf, below=1mm of m2] (m3) {M3\ \ identifier rename};
\node[leaf, below=1mm of m3] (m4) {M4\ \ wrapper extraction};
\node[leaf, below=5mm of m4] (m5) {M5\ \ sink attr obfuscation$^{\dagger}$};
\node[leaf, below=1mm of m5] (m6) {M6\ \ sink via builtins$^{\dagger}$};
\node[leaf, below=5mm of m6] (m7) {M7\ \ taint-through-dict$^{\dagger}$};

\node[fam, anchor=east] (f1) at ($(m1.west)!0.5!(m4.west) + (-9mm,0)$) {sink-\\preserving};
\node[fam, anchor=east] (f2) at ($(m5.west)!0.5!(m6.west) + (-9mm,0)$) {sink-\\targeted};
\node[fam, anchor=east] (f3) at ($(m7.west) + (-9mm,0)$) {dataflow};

\node[root, anchor=east] (r) at ($(f2.west) + (-10mm,0)$) {7\\mutators};

\draw[arr] (r.east) -- (f1.west);
\draw[arr] (r.east) -- (f2.west);
\draw[arr] (r.east) -- (f3.west);
\foreach \l in {m1,m2,m3,m4} \draw[arr] (f1.east) -- (\l.west);
\foreach \l in {m5,m6} \draw[arr] (f2.east) -- (\l.west);
\draw[arr] (f3.east) -- (m7.west);
\end{tikzpicture}
\caption{The seven adversarial mutators grouped by the detector
shortcut each family probes. The sink-preserving family alters code
around the sink; the sink-targeted family rewrites the sink call
itself; the dataflow family breaks the static lineage from a
parameter to the sink argument. Mutators marked $\dagger$ are held
out of training-time augmentation and used only as test-time
robustness probes (Section~\ref{sec:eval:model}).}
\label{fig:mutator-tax}
\end{figure}

\subsection{Design principles}
\label{sec:method:principles}

Every mutator follows three rules.

\paragraph{Semantics-preserving.} The mutated code must produce the
same runtime behavior as the original. The mutators work on the Python
AST rather than on text. A function that reads a query string and
calls \texttt{cursor.execute} on it must still read the same string
and call the same function after mutation. When the mutator cannot
prove a particular transformation is safe (for example, an identifier
rename that might collide with an imported name), it skips that node
and leaves it unchanged.

\paragraph{Targeted.} Each mutator attacks one specific shortcut. We
did not build a generic code obfuscator. Mixing the shortcuts together
would make the per-mutator F1 drop uninterpretable: a detector that
fails on a generic obfuscation tells us nothing useful, but a detector
that fails on string splitting and passes the other three is giving us
diagnostic information.

\paragraph{Realistic.} The output should look like code a real
developer might write. Identifiers are replaced with synonyms rather
than random strings. Dead code reads like ordinary scaffolding (loop
counters, unused locals) rather than obvious filler. The point is to
fool detectors that have learned the shape of typical vulnerable code,
not detectors that have learned to recognize adversarial inputs.

\subsection{Mutator 1: dead code injection}
\label{sec:method:deadcode}

The first mutator inserts no-op statements at random points inside the
function body. The example in Listing~\ref{lst:deadcode} shows the
effect on a SQL injection sample.

\begin{figure}[h]
\begin{verbatim}
# Before
def get_user(user_id):
    query = f"SELECT * FROM users WHERE id = {user_id}"
    return cursor.execute(query)

# After
def get_user(user_id):
    _unused_var = sum([])
    query = f"SELECT * FROM users WHERE id = {user_id}"
    if False:
        pass
    return cursor.execute(query)
\end{verbatim}
\caption{Dead code injection. Two statements are added; runtime
behavior is unchanged.}
\label{lst:deadcode}
\end{figure}

The shortcut this attacks is positional reasoning. A model that has
learned ``the dangerous line is usually near the top of the function''
or ``the unsafe \texttt{execute} call is two lines after the
\texttt{query} assignment'' will see those distances change after
injection.

We use three constructs: an assignment to an unused name
(\texttt{\_unused\_var = sum([])}), an \texttt{if False: pass}
statement (statically unreachable), and a \texttt{for \_ in range(0):
pass} loop (also unreachable, no iterations). All three are
side-effect free under any Python implementation, so the function's
runtime behavior cannot change.

\subsection{Mutator 2: string literal splitting}
\label{sec:method:stringsplit}

The second mutator takes a string literal and splits it into two or
more pieces joined with \texttt{+}. The runtime value is identical;
the surface text is not.

\begin{figure}[h]
\begin{verbatim}
# Before
return cursor.execute(f"SELECT * FROM users WHERE id = {uid}")

# After
return cursor.execute(
    "SEL" + "ECT * FROM" + " users WHERE id = " + f"{uid}"
)
\end{verbatim}
\caption{String literal splitting. The substring \texttt{"SELECT"} no
longer appears anywhere in the source, but the runtime SQL string is
byte-identical.}
\label{lst:stringsplit}
\end{figure}

The shortcut this attacks is substring matching. Many static analyzers
flag a file the moment it contains a literal substring like
\texttt{"SELECT "}, \texttt{"DELETE FROM"}, or \texttt{"<script>"}.
After splitting, none of those substrings exist as a contiguous token
in the source.

Three details matter for correctness. Docstrings are not split because
splitting them would mangle the output of \texttt{help()}. Type-hint
strings are not split because some libraries parse them with
\texttt{typing.get\_type\_hints} at runtime, and altering the string
content would change the parse result. F-string interpolation
expressions are not split either; only the literal text between the
\texttt{\{...\}} substitutions is eligible.

\subsection{Mutator 3: identifier rename}
\label{sec:method:rename}

The third mutator replaces local variable and parameter names with
synonyms drawn from a curated thesaurus. \texttt{user} becomes
\texttt{account}, \texttt{query} becomes \texttt{sql},
\texttt{result} becomes \texttt{outcome}. Every reference to the
renamed identifier inside the same scope is updated consistently.

\begin{figure}[h]
\begin{verbatim}
# Before
def get_user(user_id):
    query = f"SELECT * FROM users WHERE id = {user_id}"
    return cursor.execute(query)

# After
def get_user(account_pk):
    sql = f"SELECT * FROM users WHERE id = {account_pk}"
    return cursor.execute(sql)
\end{verbatim}
\caption{Identifier rename. Two locals renamed in all their
references.}
\label{lst:rename}
\end{figure}

The shortcut this attacks is identifier memorization. A classifier
trained on labeled data can latch onto specific names. If eighty
percent of the training-set SQL injection examples use a parameter
named \texttt{user\_id}, the model can shortcut on the name itself:
seeing \texttt{user\_id} flow into \texttt{execute} predicts CWE-89.
A real attacker, or any developer with a different style, might call
that parameter \texttt{pk}, \texttt{customer\_id}, or \texttt{target},
and the shortcut breaks.

Several identifiers are never renamed because doing so would change
semantics. The mutator skips \texttt{self} and \texttt{cls} (would
break method dispatch), Python builtins (\texttt{len},
\texttt{range}, \texttt{eval}, etc.), names imported from other
modules, names declared \texttt{global} or \texttt{nonlocal}, the
function's own name (would break recursion), and any token that
appears literally inside a string. The last exception matters
because string contents are data, not code: a literal
\texttt{"Looking up user\_id: \%d"} stays as written even if the
\texttt{user\_id} parameter is renamed.

The mutator also takes an intensity parameter (between 50\% and 100\%
of eligible identifiers in the function are renamed per pass). At
intensity 50\% the mutation is mild, comparable to a developer
refactor. At intensity 100\% the lexical disruption is total and
only a detector that is genuinely tracking data flow rather than
identifier patterns can survive.

\subsection{Mutator 4: wrapper extraction}
\label{sec:method:wrapper}

The fourth mutator is the most aggressive. It takes a
vulnerability-relevant function call (the sink) and hides it inside
a one-line nested helper. The outer function no longer contains a
direct reference to the sink; it contains a call to a freshly-defined
helper that in turn calls the sink.

\begin{figure}[h]
\begin{verbatim}
# Before
def get_user(user_id):
    query = f"SELECT * FROM users WHERE id = {user_id}"
    return cursor.execute(query)

# After
def get_user(user_id):
    def _wrapped_execute_138(*_a, **_kw):
        return cursor.execute(*_a, **_kw)

    query = f"SELECT * FROM users WHERE id = {user_id}"
    return _wrapped_execute_138(query)
\end{verbatim}
\caption{Wrapper extraction. The dangerous call is moved into a
one-line nested helper.}
\label{lst:wrapper}
\end{figure}

The shortcut this attacks is single-function analysis. Many detectors
work one function at a time: they enumerate the function calls
inside the function under analysis, check each callee against a
danger list, and decide based on the result. Wrapper extraction
moves the danger call inside a nested function. A detector that
only inspects the outer function sees a call to
\texttt{\_wrapped\_execute\_138}, a name not on any danger list,
and concludes the function is safe. Finding the real sink requires
following the call into the wrapper, which is interprocedural data
flow analysis. Many file-level static analyzers do not perform
this analysis, and at least some transformer architectures struggle
with it because of attention cost.

Several details ensure correctness. The wrapper uses
\texttt{*\_a, **\_kw} argument forwarding, so any call to the original
sink (positional, keyword, default, or starred-unpacking) is forwarded
unchanged. If the original call is
\texttt{await db.execute(...)}, the wrapper is generated as
\texttt{async def} with \texttt{return await ...}. The wrapper
name carries a random four-digit suffix so it cannot collide with
any existing local. If the function has a docstring, the wrapper
is inserted as \texttt{body[1]} rather than \texttt{body[0]},
keeping the docstring in its \texttt{help()}-visible position.

A small curated list of sink names guides which call to wrap when
the function contains more than one candidate. The list covers
about thirty names spanning the seven CWE classes:
\texttt{execute}, \texttt{system}, \texttt{eval}, \texttt{loads},
\texttt{open}, \texttt{get}, \texttt{Markup}, and so on. Given a
function with both \texttt{len(items)} and
\texttt{cursor.execute(query)} the mutator wraps the
security-relevant call, not the harmless one. If a function has no
eligible sink the mutator returns it unchanged.

\subsection{Mutator 5: sink attribute obfuscation}
\label{sec:method:sao}

The fifth mutator targets the sink call directly. Attribute-style
sinks of the form \texttt{value.attr(args)} are rewritten as
\texttt{getattr(value, "attr")(args)}. The two forms are
semantically equivalent under the Python attribute-lookup rules:
\texttt{getattr(x, "y")} returns the same object that
\texttt{x.y} resolves to, and applying that callable to the original
arguments produces the same return value.

\begin{figure}[h]
\begin{verbatim}
# Before
def run_cmd(user_arg):
    return os.system("ls " + user_arg)

# After
def run_cmd(user_arg):
    return getattr(os, "system")("ls " + user_arg)
\end{verbatim}
\caption{Sink attribute obfuscation. The sink identifier
\texttt{system} no longer appears as a syntactic call target; it
lives only inside a string literal.}
\label{lst:sao}
\end{figure}

The shortcut this attacks is token-level sink matching. Bandit and
Semgrep both ship rules that anchor on the textual occurrence of
sink identifiers (\texttt{B602} matches \texttt{os.system},
\texttt{B608} matches \texttt{.execute} on cursor-shaped receivers,
etc.). After the rewrite, the sink token only appears inside a
string constant, which most lexical rules do not inspect.

The mutator's target list covers attribute-style sinks across four
CWE classes: \texttt{os.system}, \texttt{os.popen}, the
\texttt{subprocess} module's call surface (CWE-78);
\texttt{.execute}, \texttt{.executemany}, \texttt{.executescript},
\texttt{.raw} on any receiver (CWE-89); the \texttt{requests} and
\texttt{httpx} HTTP verbs and \texttt{urllib(2).urlopen} (CWE-918);
and \texttt{pickle}, \texttt{cPickle}, \texttt{dill},
\texttt{yaml.unsafe\_load}, \texttt{marshal}, and \texttt{shelve}
deserialization entry points (CWE-502). Sinks where the receiver is
a runtime object (\texttt{cursor.execute},
\texttt{connection.executemany}) are handled with a wildcard
receiver in the target list, since the attribute name is what makes
the sink dangerous, not the import-time module. The full target
list is in \texttt{src/red\_team/mutators/sink\_attr\_obfuscate.py}.

CWE-94 (\texttt{eval}, \texttt{exec}, \texttt{compile},
\texttt{\_\_import\_\_}) is a bare-builtin call rather than an
attribute access, and CWE-22 path-traversal and CWE-79 XSS sinks
are framework-specific calls that do not fit the
\texttt{getattr} rewrite cleanly. CWE-94 is the target of Mutator 6;
the remaining two CWE classes are not yet exercised by a
sink-targeted mutator and are listed as a residual gap in
Section~\ref{sec:threats:construct}.

The mutator is held out of training-time augmentation, alongside
\texttt{string\_split} and \texttt{sink\_via\_globals}.

\subsection{Mutator 6: sink via globals}
\label{sec:method:svg}

The sixth mutator is the bare-builtin counterpart to M5. Bare
builtin sinks of the form \texttt{eval(arg)} are rewritten as
\texttt{\_\_import\_\_("builtins")} \texttt{.\_\_dict\_\_["eval"](arg)}.
The two forms are runtime-identical: the
\texttt{\_\_import\_\_("builtins")} call returns the same module
object that supplies the bare name \texttt{eval} at lookup time, and
its \texttt{\_\_dict\_\_} attribute exposes the builtin under the
same key.

\begin{figure}[h]
\begin{verbatim}
# Before
def render_template(spec):
    return eval(spec)

# After
def render_template(spec):
    return __import__("builtins").__dict__["eval"](spec)
\end{verbatim}
\caption{Sink-via-globals rewrite for a CWE-94 \texttt{eval} sink.
The sink token only appears as a string literal inside a
\texttt{\_\_dict\_\_} subscript.}
\label{lst:svg}
\end{figure}

The shortcut this attacks is the same token-level sink match that
M5 attacks, exercised on a different syntactic position: M5 covers
\texttt{ast.Attribute} sinks, M6 covers \texttt{ast.Name} sinks. We
keep them as separate mutators rather than as a single mutator that
covers both positions because they apply to disjoint CWE classes
(M5 covers four CWEs; M6 covers one, CWE-94) and reporting their
robustness drops separately preserves that information.

The mutator's target set is the four CWE-94 bare builtins
\texttt{eval}, \texttt{exec}, \texttt{compile}, and
\texttt{\_\_import\_\_}. We chose \texttt{\_\_dict\_\_} indexing
over \texttt{getattr}-style indirection deliberately: M5 already
uses \texttt{getattr}, and a detector that learns to flag
\texttt{getattr(\_, "eval")} would still miss
\texttt{\_\_dict\_\_["eval"]}. Routing through a different
indirection mechanism gives a distinct obfuscation axis. We did not
use the naive \texttt{globals()["eval"](\ldots)} form despite the
mutator name: bare builtins are not stored in the module's
\texttt{globals()} dict, so the naive rewrite raises
\texttt{KeyError} at runtime; \texttt{\_\_import\_\_("builtins").%
\_\_dict\_\_["eval"]} is the semantically faithful counterpart.

The mutator is held out of training-time augmentation alongside M2
and M5.

\subsection{Mutator 7: taint through dict}
\label{sec:method:ttd}

The seventh mutator targets a third shortcut, orthogonal to the
sink-preserving and sink-targeted families. M1--M6 either change the
code around the sink or change the sink identifier; M7 leaves both
intact and instead breaks the \emph{flow} between a tainted source
and the sink call. Every sink call's first positional argument is
routed through a one-entry dict literal: \texttt{f(expr)} becomes
\texttt{f(\{"\_a": expr\}["\_a"])}. The wrapped expression is
constructed and immediately re-extracted, so the value reaching the
sink is bitwise identical to the original argument.

\begin{figure}[h]
\begin{verbatim}
# Before
def get_user(user_id):
    return cursor.execute(
        "SELECT * FROM users WHERE id = " + user_id
    )

# After
def get_user(user_id):
    return cursor.execute(
        {"_a": "SELECT * FROM users WHERE id = " + user_id}["_a"]
    )
\end{verbatim}
\caption{Taint-through-dict rewrite. The sink identifier
\texttt{cursor.execute} is unchanged and the sink argument
evaluates to the same string; the static lineage from the
\texttt{user\_id} parameter to the sink argument now passes
through a dict construction and a subscript.}
\label{lst:ttd}
\end{figure}

The shortcut this attacks is taint-flow tracking that does not
propagate through dict reads. A pattern-matching SAST tool that
flags any \texttt{cursor.execute(<expr>)} call sees the sink and
fires regardless of how the argument was assembled; we therefore
expect Bandit and the non-taint Semgrep configurations to be
invariant to this mutator. A dataflow-aware detector that follows
\texttt{user\_id} forward, however, must keep the taint label
through the \texttt{\{"\_a": user\_id\}} dict construction and
the subsequent \texttt{["\_a"]} subscript to flag the call; a
detector that drops taint at the dict boundary will lose the
sink. The mutator is included to (i) exercise this third axis in
the benchmark, (ii) provide a probe that will distinguish
dataflow-aware detectors from pure pattern-matching tools when
those detectors (CodeQL, taint-mode Semgrep configurations) are
added in future work.

The mutator's target sinks are the union of the M5 and M6 sets:
attribute-style sinks across CWE-78/89/918/502 plus bare-builtin
sinks across CWE-94. Sinks whose first positional argument is a
literal constant are skipped (there is no taint to flow), as are
sinks called with kwargs only. The mutator is held out of
training-time augmentation alongside M2, M5, and M6.

\subsection{Composition}
\label{sec:method:composition}

At evaluation time we apply between two and three mutators to each
test sample. Two is the lower bound because a single mutator probes
one shortcut at a time, and the aggregate robustness number is more
informative when at least two shortcuts are in play. Three is the
upper bound because applying every mutator at once over-mangles the
code: even valid mutations stop being instructive, and the result
is no longer plausibly a thing a real developer would write.

Mutator selection and application order are randomized per sample
but seeded by the sample identifier, so re-running the evaluation
produces identical adversarial variants. The seed is published with
the dataset.

For the headline robustness table (Section~\ref{sec:evaluation}) we
also report each mutator applied in isolation, producing seven
single-mutator test sets in addition to the composed one. The
single-mutator numbers tell us which shortcut breaks which tool: a
tool whose F1 collapses on \texttt{sink\_attr\_obfuscate} but holds
steady on the four sink-preserving mutators is doing token-level
sink matching, and the per-mutator columns of
Table~\ref{tab:robustness} make that visible.

The three module-scope mutators (M5, M6, and M7) operate on the
whole module's AST rather than the first function alone, because
the sinks they target can live anywhere (helper functions, class
methods, module-level statements). M1--M4 mutate only the first
function in each module under the standard per-function pipeline.
When the composed variant draws from all seven mutators through
that pipeline, M5--M7 are scoped to the first function rather
than the whole module, which weakens their effect in the composed
condition. The single-mutator M5, M6, and M7 variants apply the
full module-scope rewrite and are the authoritative source for the
sink-targeted and dataflow robustness numbers reported in
Section~\ref{sec:eval:robustness}.

\subsection{Semantic equivalence}
\label{sec:method:equivalence}

The argument that mutations preserve behavior rests on AST-level
invariants. Two checks back this up.

First, every mutated sample is required to parse: a mutation that
produces unparseable source is rejected and the sample is left
unchanged in the evaluation set. This catches construction bugs in
the mutators themselves rather than letting them silently corrupt
the test set.

Second, for each mutator the implementation enumerates the AST node
types it can touch and the transformations it can apply to each.
Those transformations are individually provably semantics-preserving
under the Python evaluation rules: \texttt{if False: pass} is
statically unreachable; argument forwarding with
\texttt{*\_a, **\_kw} is the documented forwarding contract;
splitting a string literal with \texttt{+} produces an equal string.
We do not perform dynamic equivalence testing (running the original
and the mutated code to compare their outputs) because most samples
are fragments of larger applications and cannot be executed
standalone.

A residual concern is reflection and metaprogramming. A function
that uses \texttt{locals()} introspection or
\texttt{inspect.getsource} to read its own variable names will
behave differently after identifier rename. A function whose
control flow depends on the source-level presence of a specific
substring (vanishingly rare in practice) will behave differently
after string splitting. The audit samples we reviewed contained no
such cases, but downstream users who care can filter the test set
with a static metaprogramming check.
