\section{Threats to Validity}
\label{sec:threats}

We discuss threats to validity along the categories of Wohlin
et al.~\cite{wohlin}. Each category lists the principal risks
specific to this benchmark, the mitigations applied during construction,
and the residual exposure that downstream consumers of the dataset and
the evaluation should be aware of.

\subsection{Construct validity}
\label{sec:threats:construct}

The principal construct-validity risk is that the labels we treat as
ground truth do not in fact identify the line on which the vulnerability
is exploitable. We mitigate this with the three-layer methodology of
Section~\ref{sec:dataset}, which combines a sink-presence filter, three
rounds of independent adversarial audit, and a diff-based filter that
requires the sink line to have been modified between
\texttt{code\_before} and \texttt{code\_after}. The audited residual
false-positive rate is 26\%, down from 52\% on the unfiltered CVE
labels. The 26\% figure is reported in
Section~\ref{sec:dataset:diff} and is a property of the dataset rather
than of any single detector. We do not claim a noise-free label set.
Two practical consequences follow. First, an absolute accuracy reported
on this benchmark inherits a 26\% audit residual and should not be read
as a tool's true accuracy on a hypothetical noise-free corpus. Second,
the comparison \emph{between} two tools on the same labels is robust to
this residual because both tools score against the same noisy ground
truth.

A second construct-validity concern is that the sink-presence filter is
itself a definition of what counts as a correct label. The sink regex
set, the comment-stripping rules, and the proximity gates for CWE-22
and CWE-798 (during its inclusion period) collectively determine which
samples are admitted to the benchmark. A detector that uses the same
sink regexes inherits a measurement advantage. The filter source code,
the full pattern set, and the audit reports for every round are
released alongside the dataset so that this dependence is auditable
rather than hidden. We additionally report the rejection manifest with
per-sample reason codes for every dropped sample, enabling a downstream
reviewer to refute or replicate the filter's decisions.

A third construct-validity concern is the strength of the
adversarial mutator suite (Section~\ref{sec:methodology}). All six
mutators are semantics-preserving by AST-level construction. They
divide into two families with sharply different effects on
detectors. The \emph{sink-preserving} family —
\texttt{dead\_code\_injection}, \texttt{variable\_rename},
\texttt{wrapper\_extraction}, and the held-out \texttt{string\_split}
— modifies code around the sink (unreachable statements, identifier
substitution, helper-function hoisting, string fragmentation) but
preserves the sink token verbatim. All three detector families are
invariant to this family ($\Delta$F1 between $-0.04$ and $+0.01$),
and \texttt{string\_split} produces a change indistinguishable from
the three training-augmented mutators on the learned detector,
evidence that the four sink-preserving mutators occupy a common
transformation axis rather than four independent ones. The
\emph{sink-targeted} family — \texttt{sink\_attr\_obfuscate} (SAO)
and \texttt{sink\_via\_globals} (SVG), both held out of training-time
augmentation — rewrites the sink itself: SAO via
\texttt{getattr}-mediated attribute access for CWE-78/89/918/502
sinks, SVG via \texttt{\_\_dict\_\_} indexing on the builtins module
for CWE-94 sinks. On this family the two SAST tools show a
robustness cliff (Bandit $\Delta$F1 $+0.187$/$+0.143$, Semgrep
$+0.109$/$+0.132$; the corresponding SWR drops range 0.196--0.347)
while the learned detector remains invariant ($\Delta$F1 between
$-0.04$ and $0.00$). The cliff replicates across two mutators
attacking disjoint sink classes, which we read as evidence that the
sink-token axis is a property of pattern-matching SAST itself rather
than an artifact of a single rewriting trick. A third mutator,
\texttt{taint\_through\_dict} (TTD), exercises a separate
\emph{dataflow} axis: the sink identifier and argument value are
preserved but the static lineage from a parameter to the sink
argument now passes through a one-entry dict construction and
subscript. TTD is invariant on every detector reported here
($|\Delta\textrm{F1}| \le 0.030$, $|\Delta\textrm{SWR}| \le 0.029$);
we read the null result as a property of the current detector set
(none performs dataflow analysis) rather than a property of the
probe, and we retain TTD in the benchmark as the slot for future
dataflow-aware detectors (Section~\ref{sec:eval:robustness}).

Four residual concerns remain. First, one further sink-targeted
axis is not yet exercised: dynamic dispatch through receiver
indirection (\texttt{getattr(cursor, attr\_name)(query)} where
\texttt{attr\_name} is a runtime variable). It could reveal an
independent axis on which the learned detector is not invariant.
Second, the SVG family applies to only 6 of the 67 test positives
(all CWE-94, all using \texttt{eval}); its drop numbers are noisier
than the SAO numbers and we rely on the agreement with the SAO
direction as cross-validation rather than treating SVG alone as
conclusive. Third, TTD's null result is uninformative about
detector dataflow capability because none of the three current
detectors performs dataflow analysis; the probe is a measurement
slot rather than a measurement, and the conclusion ``no detector
breaks on dataflow obfuscation'' should be read as ``no detector
in the current set has dataflow tracking to break.'' Fourth, the
mutator suite is evaluated against a single trained model
(Section~\ref{sec:eval:model}); the conclusion that sink-targeted
obfuscation does not break a learned detector should be read as a
property of this specific GraphCodeBERT checkpoint and not as a
universal property of learned detectors.

\subsection{Internal validity}
\label{sec:threats:internal}

Single-CVE clustering is the principal internal-validity risk: a single
high-impact CVE typically generates several near-identical sample files,
which can dominate per-class metrics if they fall on the same side of
the split. We mitigate this with a per-CVE sample cap of three at the
scraping stage and with a repo-group constraint in the stratified
splitter (Section~\ref{sec:dataset:final}), which forces all samples
that share a repository to be assigned to the same train, validation,
or test split. We verify zero cross-split repository leakage after the
split is constructed; the verification is part of the released
build-pipeline tests.

The evaluation uses a single seed for the splitter (seed=42) and a
single run per detector. We do not report multi-seed mean and standard
deviation. For the static analysis detectors this is sound because
their output is deterministic given the input and the rule set. For
the LLM and trained-model detectors that will be added in the
evaluation phase this is a residual risk, and the planned multi-seed
protocol is documented as future work
(Section~\ref{sec:conclusion}). Detector hyperparameters are pinned to
their default values for the static tools (Bandit~1.9.4, Semgrep~1.163.0
with the three published rulesets named in
Section~\ref{sec:eval:setup}); for the LLM and trained-model
detectors that will be added later, hyperparameters and prompts will be
fixed and released with the evaluation script.

\subsection{External validity}
\label{sec:threats:external}

The benchmark is Python-specific. The label-validation methodology
generalizes (the sink-presence filter, the audit cycle, and the
diff-based filter are language-agnostic in design), but the sink
pattern set we ship is calibrated for Python and would need to be
re-derived for another target language. We do not claim that the
detector rankings reported here transfer to other languages.

The benchmark targets seven sink-shaped CWE classes drawn from the
MITRE Top-25. Structural CWEs (CWE-352, CWE-862, CWE-863, CWE-306,
CWE-284, CWE-639, CWE-200) are excluded on methodological grounds, as
documented in Section~\ref{sec:dataset:scope}. A detector that handles
structural CWEs well is not measurable on this benchmark, and the
scores reported here are not a comprehensive Python vulnerability
detection ranking.

The scraped sample set is biased by the repositories present in CVEfixes,
GHSA Advisory DB, OSV, PyPA, and NVD-targeted scrapes. Popular Python
projects (Django, Flask, FastAPI applications) are over-represented;
small or unmaintained projects with security relevant code are
under-represented. The hand-curated canonical samples partially offset
this by introducing textbook examples that no real repository may
contain in idiomatic form, but they account for only 20 of the 871
positives.

Finally, the 132-sample test set is small. Per-class test sizes range
from 3 (CWE-22) to 32 (CWE-89). For the smallest classes the per-class
F1 carries a wide confidence interval, which we flag in
Table~\ref{tab:per-cwe-counts} and the per-class results table. A
follow-up release will increase the per-class sample count for
CWE-22 and CWE-918, where the current size is the dominant source of
per-class measurement noise.

\subsection{Conclusion validity}
\label{sec:threats:conclusion}

The evaluation reports point estimates of detector metrics rather than
confidence intervals. For the static analysis baseline this is adequate
because the rule-based detectors are deterministic and the per-class
counts are exhaustively enumerated. The trained-model row reports a
single run with a fixed seed and is deterministic given the seed and
training data; the protocol does not yet include multi-seed
re-training. For the LLM rows that will be added later, point estimates
without intervals would be inappropriate; the evaluation protocol for
those rows requires multiple runs and bootstrap confidence intervals on
the headline metric, and that protocol is documented in
Section~\ref{sec:conclusion}.

A related concern is the size of the test set. The 67-positive
mutator-eligible subset produces a per-sample F1 swing of approximately
1.5\%, so an absolute change in macro-F1 below 0.04 lies near the
noise floor of this benchmark and should not be read as a signed
robustness result. The $-0.036$ mean drop reported for the trained
detector in Section~\ref{sec:eval:robustness} is at the boundary of
this noise floor and is reported as evidence of invariance rather
than of improvement under mutation. Per-class confidence intervals
are particularly wide for CWE-22, CWE-78, and CWE-918, where the
positive test count is below seven and a single prediction flip
changes F1 by 14 to 33 points.

We do not perform a paired statistical test between detectors on the
current evaluation. Once the LLM rows are available, a paired bootstrap
test (resampling samples rather than runs) is the planned protocol
for between-detector comparisons.

The evaluation in Section~\ref{sec:evaluation} reports a five-measure
profile (macro-F1, detection MCC, Severity-Weighted Recall,
hierarchical CWE macro-accuracy, weighted Cohen's $\kappa$) rather
than the CASTLE Score~\cite{castle} or a security-weighted F-measure
such as F3. The macro-F1 column is kept for continuity with prior
Python evaluations that report it. MCC~\cite{chicco2020mcc} replaces
accuracy on the binary vulnerable-versus-safe collapse because it is
chance-corrected and stays informative under the rough 50/50 prevalence
of the clean test set. Severity-Weighted Recall replaces an
ordinal-rank severity bonus with a CVSS-base-score-weighted recall on
the subset of positives that ship a CVSS score in the advisory feed
(27 of the 67 positives in the current test set). HCMA replaces
flat exact-match accuracy with the Wu--Palmer
similarity~\cite{wupalmer1994} between the detector's prediction and
the ground-truth CWE on a hand-coded subtree of MITRE CWE 4.14
(\texttt{src/eval/cwe\_map.py}), so that close-sibling
mis-classifications (e.g. predicting CWE-94 when truth is CWE-78,
both under CWE-74 Injection) receive partial credit. Weighted
Cohen's $\kappa$ (Cohen~\cite{cohen1968kappa}) is a chance-corrected
rater-agreement statistic that uses $1{-}$Wu--Palmer as its
disagreement weight; it gives a single bounded headline number with
partial credit for near-misses, at the cost of the well-documented
$\kappa$ paradox under highly skewed prevalence. We report raw
observed agreement alongside $\kappa$ as the methodology requires.
Pairwise bootstrap intervals on $\kappa$ are deferred to the LLM
extension of this evaluation.

\subsection{Summary of limitations}
\label{sec:threats:limitations}

We close with a consolidated summary of the two principal limitations
that reviewers should weigh against the headline results. Both are
discussed in detail in the subsections above; we surface them here
because either alone bounds the conclusions a reader can draw from
this work.

\paragraph{Test-set size.} The 132-sample test set
(67 positives plus 65 hard negatives, distributed across seven CWE
classes) is small by the standards of post-PrimeVul vulnerability
benchmarks: PrimeVul~\cite{primevul} ships approximately 7{,}000
vulnerable samples, and RealVuln~\cite{realvuln} reports 796.
Per-class F1 confidence intervals are correspondingly wide,
particularly for CWE-22, CWE-78, and CWE-918, where the positive
count is below seven and a single prediction flip changes F1 by
14--33 points (Section~\ref{sec:threats:external},
Section~\ref{sec:threats:conclusion}). The
\texttt{sink\_via\_globals} mutator additionally applies to only 6
of the 67 positives (all CWE-94 \texttt{eval} sinks); its drop
numbers should be read as direction-confirming the
\texttt{sink\_attr\_obfuscate} cliff rather than as a
high-confidence point estimate
(Section~\ref{sec:threats:construct}).

\paragraph{Single-model learned evaluation.} The learned-detector
robustness results in Section~\ref{sec:eval:model} are produced by a
single fine-tuned GraphCodeBERT~\cite{graphcodebert} checkpoint. The
invariance of this detector to both mutator families is a property
of this checkpoint and not a universal property of learned
vulnerability detectors. A multi-model evaluation (a second
Transformer encoder and a non-Transformer baseline) is the natural
extension and is left as a planned follow-up. Until that evaluation
is run, the claim that learned detectors generalize across the
sink-targeted mutator axis is scoped to this specific detector.
